### 2023-09-19 assignment_03stat405.qmd
Add files via upload

---
title: 'Assignment 3 STAT 405'
format:
  pdf: default
pdf-engine: pdflatex  # You may need to remove this line
---

* Please produce your assignment as a pdf (knit to pdf).
See instructions in announcements if you have not downloaded a LaTeX distributor.

* Please also submit your
qmd file (it will not be graded but we want it for reference
purposes).
* Please use the provided qmd file as a basis for your first assignment
(do not submit an R script file), putting your solutions
inside the empty blocks that follow "Solution:"
* You will have to produce your future qmd files.
* 20 points per exercise

<br />

#### Exercise 1:
Using paste (or paste0), create the following Latex formula: a_{1}X_{1}ˆ0 + a_{2}X_{2}ˆ1+ a_{3}X_{3}ˆ2 + a_{4}X_{4}ˆ3 + a_{5}X_{5}ˆ4

Solution:
```{r}
paste0("a_{",1:5,"}X_{",1:5,"}^",0:4, collapse = "+")
```

<br />

#### Exercise 2:
A bucket contains 12 rocks: five yellow rocks and seven black rocks. If we select three rocks at random, what is the probability that exactly one yellow rock gets selected? 
```{r}
#1. (9 points) using a for loop with an if() clause and accumulator (Hint: The strategy is to sample 3 rocks without replacement from sample_space
sample_space <- c(rep("Yellow", 5), rep("Blue", 7))
n <- 10000
set.seed(123)
accumulator <- 0
for(i in 1:n){
  curr_sample <- sample(sample_space, 3)
  if( "Yellow" %in% curr_sample && table(curr_sample)[["Yellow"]] == 1 )
    accumulator <- accumulator +1
}
accumulator/n

#2 (8 points) using a for loop with an ifelse() function and accumulator
set.seed(123)
accumulator <- 0
for(i in 1:n){
  curr_sample <- sample(sample_space, 3)
  ifelse( "Yellow" %in% curr_sample && table(curr_sample)[["Yellow"]] == 1,  accumulator <- accumulator +1, next)
}
accumulator/n

#3. (8 points) Vectorial approach using sapply (no loops, no if() or ifelse(), no accumulator).
set.seed(123)
simulated_results <- sapply(1:n, function(x) {
  curr_sample <- sample(sample_space, 3)
  num_yellow_selected <- sum(curr_sample == "Yellow")
  return(num_yellow_selected)
})

simulated_probability <- sum(simulated_results == 1) / n
simulated_probability
```

<br />

#### Exercise 3:
1. (12.5 points) In the previous exercise, what is the probability that at most one yellow rock gets selected?
2. (12.5 points) In the previous exercise, what is the probability that at least one yellow rock gets selected?

Solution:
```{r}
#1
set.seed(123)
simulated_results <- sapply(1:n, function(x) {
  curr_sample <- sample(sample_space, 3)
  num_yellow_selected <- sum(curr_sample == "Yellow")
  return(num_yellow_selected)
})

simulated_probability <- sum(simulated_results == 1 | simulated_results == 0) / n
simulated_probability

#2

set.seed(123)
simulated_results <- sapply(1:n, function(x) {
  curr_sample <- sample(sample_space, 3)
  num_yellow_selected <- sum(curr_sample == "Yellow")
  return(num_yellow_selected)
})

simulated_probability <- sum(simulated_results == 1 | simulated_results == 2 | simulated_results == 3) / n
simulated_probability

```
#### Exercise 4:

1. (9 points) Suppose you take $1000 to the casino and you play roulette. Every turn, you decide to bet 100 dollars on black. If you win, which occurs with probability 9/19, you win 200, which is a net profit of 100. If you lose, you get nothing. Using a while loop, write code to simulate your adventure until you lose all your money. How many rounds will it take?
2. (8 points) Now, repeat the above simulation 1,000 times. What is the average number of times you get to play for each $1000 bankroll you use? Plot a histogram of the number of times you get to play for each bankroll. Comment a sentence on the histogram
3. (8 points) Suppose you quit if you double your money. Estimate the probability you double your money by running the simulation 1,000 times.
```{r}
#1
set.seed(123)
play_roulette <- function(x) {
money <- 1000
rounds <- 0
outcomes <- c("win", "lose")
  while(money >0){
  rounds<- rounds + 1
  money <- money - 100
  play_outcome<- sample(outcomes, size = 1, prob=c(9/19, 10/19))
  if(play_outcome == "win"){
    money <- money + 200
    }
  }
return(rounds)
}

print(play_roulette())
#2
set.seed(123)
num_sim <- 1000
roulette_sim <- sapply(1:num_sim, play_roulette)
mean(roulette_sim)
hist(roulette_sim, main = "Histogram of Roulette Simulation", xlab = "Rounds")

#This histogram shows that most rounds will be in the lower range of possible 
#rounds/is right skewed.Very unlikely to play a lot of rounds.

#3
set.seed(123)
play_double_or_nothing_roulette <- function(x) {
money <- 1000
rounds <- 0
outcomes <- c("win", "lose")
  while(money >0 && money < 2000){
    rounds<- rounds + 1
    money <- money - 100
    play_outcome<- sample(outcomes, size = 1, prob=c(9/19, 10/19))
    if(play_outcome == "win"){
      money <- money + 200
    }
  }
  if(money>= 2000){
    return(1)
  }
return(0)
}
num_sim <- 1000
double_sim <- sapply(1:num_sim, play_double_or_nothing_roulette)
#Probability of doubling your money
sum(double_sim)/1000

```

```

<br />

### 2023-09-19 STAT405Project.qmd
Changed name
Formatted Document

--- 2023-09-19 assignment_03stat405.qmd
@@
 ---
-title: 'Assignment 3 STAT 405'
+title: 'STAT 405 Project'
 format:
@@
 
-* Please produce your assignment as a pdf (knit to pdf).
-See instructions in announcements if you have not downloaded a LaTeX distributor.
-
-* Please also submit your
-qmd file (it will not be graded but we want it for reference
-purposes).
-* Please use the provided qmd file as a basis for your first assignment
-(do not submit an R script file), putting your solutions
-inside the empty blocks that follow "Solution:"
-* You will have to produce your future qmd files.
-* 20 points per exercise
-
 <br />
-
-#### Exercise 1:
-Using paste (or paste0), create the following Latex formula: a_{1}X_{1}ˆ0 + a_{2}X_{2}ˆ1+ a_{3}X_{3}ˆ2 + a_{4}X_{4}ˆ3 + a_{5}X_{5}ˆ4
 
@@
 ```{r}
-paste0("a_{",1:5,"}X_{",1:5,"}^",0:4, collapse = "+")
+
 ```
@@
 
-#### Exercise 2:
-A bucket contains 12 rocks: five yellow rocks and seven black rocks. If we select three rocks at random, what is the probability that exactly one yellow rock gets selected? 
-```{r}
-#1. (9 points) using a for loop with an if() clause and accumulator (Hint: The strategy is to sample 3 rocks without replacement from sample_space
-sample_space <- c(rep("Yellow", 5), rep("Blue", 7))
-n <- 10000
-set.seed(123)
-accumulator <- 0
-for(i in 1:n){
-  curr_sample <- sample(sample_space, 3)
-  if( "Yellow" %in% curr_sample && table(curr_sample)[["Yellow"]] == 1 )
-    accumulator <- accumulator +1
-}
-accumulator/n
-
-#2 (8 points) using a for loop with an ifelse() function and accumulator
-set.seed(123)
-accumulator <- 0
-for(i in 1:n){
-  curr_sample <- sample(sample_space, 3)
-  ifelse( "Yellow" %in% curr_sample && table(curr_sample)[["Yellow"]] == 1,  accumulator <- accumulator +1, next)
-}
-accumulator/n
-
-#3. (8 points) Vectorial approach using sapply (no loops, no if() or ifelse(), no accumulator).
-set.seed(123)
-simulated_results <- sapply(1:n, function(x) {
-  curr_sample <- sample(sample_space, 3)
-  num_yellow_selected <- sum(curr_sample == "Yellow")
-  return(num_yellow_selected)
-})
-
-simulated_probability <- sum(simulated_results == 1) / n
-simulated_probability
-```
-
 <br />
-
-#### Exercise 3:
-1. (12.5 points) In the previous exercise, what is the probability that at most one yellow rock gets selected?
-2. (12.5 points) In the previous exercise, what is the probability that at least one yellow rock gets selected?
-
-Solution:
-```{r}
-#1
-set.seed(123)
-simulated_results <- sapply(1:n, function(x) {
-  curr_sample <- sample(sample_space, 3)
-  num_yellow_selected <- sum(curr_sample == "Yellow")
-  return(num_yellow_selected)
-})
-
-simulated_probability <- sum(simulated_results == 1 | simulated_results == 0) / n
-simulated_probability
-
-#2
-
-set.seed(123)
-simulated_results <- sapply(1:n, function(x) {
-  curr_sample <- sample(sample_space, 3)
-  num_yellow_selected <- sum(curr_sample == "Yellow")
-  return(num_yellow_selected)
-})
-
-simulated_probability <- sum(simulated_results == 1 | simulated_results == 2 | simulated_results == 3) / n
-simulated_probability
-
-```
-#### Exercise 4:
-
-1.
-(9 points) Suppose you take $1000 to the casino and you play roulette.
-Every turn, you decide to bet 100 dollars on black.
-If you win, which occurs with probability 9/19, you win 200, which is a net profit of 100.
-If you lose, you get nothing.
-Using a while loop, write code to simulate your adventure until you lose all your money.
-How many rounds will it take?
-2.
-(8 points) Now, repeat the above simulation 1,000 times.
-What is the average number of times you get to play for each $1000 bankroll you use?
-Plot a histogram of the number of times you get to play for each bankroll.
-Comment a sentence on the histogram
-3. (8 points) Suppose you quit if you double your money. Estimate the probability you double your money by running the simulation 1,000 times.
-```{r}
-#1
-set.seed(123)
-play_roulette <- function(x) {
-money <- 1000
-rounds <- 0
-outcomes <- c("win", "lose")
-  while(money >0){
-  rounds<- rounds + 1
-  money <- money - 100
-  play_outcome<- sample(outcomes, size = 1, prob=c(9/19, 10/19))
-  if(play_outcome == "win"){
-    money <- money + 200
-    }
-  }
-return(rounds)
-}
-
-print(play_roulette())
-#2
-set.seed(123)
-num_sim <- 1000
-roulette_sim <- sapply(1:num_sim, play_roulette)
-mean(roulette_sim)
-hist(roulette_sim, main = "Histogram of Roulette Simulation", xlab = "Rounds")
-
-#This histogram shows that most rounds will be in the lower range of possible 
-#rounds/is right skewed.Very unlikely to play a lot of rounds.
-
-#3
-set.seed(123)
-play_double_or_nothing_roulette <- function(x) {
-money <- 1000
-rounds <- 0
-outcomes <- c("win", "lose")
-  while(money >0 && money < 2000){
-    rounds<- rounds + 1
-    money <- money - 100
-    play_outcome<- sample(outcomes, size = 1, prob=c(9/19, 10/19))
-    if(play_outcome == "win"){
-      money <- money + 200
-    }
-  }
-  if(money>= 2000){
-    return(1)
-  }
-return(0)
-}
-num_sim <- 1000
-double_sim <- sapply(1:num_sim, play_double_or_nothing_roulette)
-#Probability of doubling your money
-sum(double_sim)/1000
-
-```
-
-```
-
-<br />

### 2023-09-21 STAT405Project.qmd
fuck
Formatting
cleaned data
Added plots

@@
   pdf: default
+
 pdf-engine: pdflatex  # You may need to remove this line
+echo: false
 ---
@@
 
-Solution:
 ```{r}
+#install.packages("readr")
+library(readr)
+
+data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)
+
+cleaned_data <- na.omit(data_set)
 
@@
 
+```{r}
+y <- cleaned_data$`Total Charges`
+```
+
+```{r}
+plot(length_of_stay, y,  xlab = "Length of Stay", main = "Price vs Length of Stay", ylab = "Price")
+
+lmfs <- lm(y ~ length_of_stay)
+abline(lmfs, col = "RED")
+```
+
+```{r}
+length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
+boxplot(cleaned_data$'Total Costs', 
+        main="Box Plot of Total Costs",
+        ylab= "Total costs")
+```
+
+```{r}
+options(repr.plot.width = 100, repr.plot.height = 100)
+
+cleaned_data$`Payment Typology 1` <- as.factor(cleaned_data$`Payment Typology 1`)
+payments <- table(cleaned_data$`Payment Typology 1`)
+pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA)
+legend("left", legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
+```
+
+```{r}
+admission_freq <- table(cleaned_data$`Type of Admission`)
+
+barplot(admission_freq, xlab = "Type Of Admission", main = "Frequencies for Type of Admission", ylab = "Frequency")
+```
+```{r}
+medianCosts <- tapply(cleaned_data$`Total Costs`, cleaned_data$Race, median)
+barplot(medianCosts, main = "Median Cost by Race", xlab = "Race", ylab = "Median Cost", col = "purple", names.arg = sort(names(medianCosts)), ylim = c(0, max(medianCosts) * 1.5))
+```
+```{r}
+ageGroups <- table(cleaned_data$`Age Group`)
+barplot(ageGroups, main = "Age Group Distribution", xlab = "Age Group", 
+        ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)) 
+```
+
 <br />

### 2023-09-23 STAT405Project.qmd
Changed plots

@@
 ```{r}
+length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
 plot(length_of_stay, y,  xlab = "Length of Stay", main = "Price vs Length of Stay", ylab = "Price")
@@
 ```{r}
-length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
+
 boxplot(cleaned_data$'Total Costs', 
@@
 ```{r}
-admission_freq <- table(cleaned_data$`Type of Admission`)
 
-barplot(admission_freq, xlab = "Type Of Admission", main = "Frequencies for Type of Admission", ylab = "Frequency")
+# Create the line chart
+plot(cleaned_data$`Length of Stay`, cleaned_data$`Total Charges`, type = "l", col = "blue", xlab = "Length of Stay", ylab = "Total Charges", main = "Trends in Total Charges and Length of Stay")
+
+# Add a line for Total Costs
+lines(cleaned_data$`Birth Weight`, cleaned_data$`Total Costs`, col = "red")
+
+# Add a legend
+legend("topright", legend = c("Total Charges", "Total Costs"), col = c("blue", "red"), lty = 1)
+
+# Add a grid
+grid()
+
+#Total cost doesn't change even if length of stay increases and total charge increases.
+
 ```
 ```{r}
-medianCosts <- tapply(cleaned_data$`Total Costs`, cleaned_data$Race, median)
-barplot(medianCosts, main = "Median Cost by Race", xlab = "Race", ylab = "Median Cost", col = "purple", names.arg = sort(names(medianCosts)), ylim = c(0, max(medianCosts) * 1.5))
+
+heatmap_data <- table(cleaned_data$`Age Group`, cleaned_data$Gender)
+
+row_names <- as.factor(rownames(heatmap_data))
+col_names <- as.factor(colnames(heatmap_data))
+
+heatmap(heatmap_data,
+        Rowv = NA,          
+        Colv = NA,       
+        col = colorRampPalette(c("white", "blue"))(50),  
+        xlab = "Gender",    
+        main = "Gender v. Age Group",  
+        labRow = row_names, 
+        labCol = col_names)
+
 ```

### 2023-09-23 STAT405Project.qmd
added titles

@@
 ---
-title: 'STAT 405 Project'
-format:
-  pdf: default
+title: "Healthcare Insights and Sociodemographic Analysis of New York: A Comprehensive Analysis of Patient Records in 2013"
+format: pdf
+author: Abel Limachi, Austin Tran, Joel Villarino, and Krish Kumar (Group 3)
+editor: visual
+---
+##
 
-pdf-engine: pdflatex  # You may need to remove this line
-echo: false
----
+*(last updated 09/21/2023)*
+
+**Introduction:**
+
+In the context of the ever-evolving landscape of healthcare, our choice of dataset stems from the pressing need to delve into an in-depth analysis of the healthcare dynamics prevailing in today's society.
+New York, being one of the most densely populated states in the United States, presents a unique microcosm of diverse populations and a wide spectrum of urban and rural healthcare challenges.
+Motivated by the significance of understanding these complexities, our research group has selected a comprehensive dataset encompassing 2,428,905 patient records, representing healthcare interactions across various hospitals throughout New York in the year 2013.
+
+This rich dataset spans a spectrum of information, including patient demographics, detailed medical records, financial aspects related to their hospitalization, and critical contextual information pertinent to each patient's stay.
+This extensive dataset holds immense potential for generating compelling visualizations and conducting statistical analyses, offering valuable insights into the intricacies of the New York healthcare system while also providing broader insights that can contribute to the understanding of healthcare systems more generally.
+
+The following project contains different visualizations and statistics that will provide you more information on the different social and economic factors associated with patients in the state of New York in 2013.
+
+**Dataset:**
+
+2013 Hospital Inpatient Discharges in the State of New York
+
+*The following data set was received from the official New York State Health website on heath.data.ny.gov*
+
+<https://health.data.ny.gov/Health/Hospital-Inpatient-Discharges-SPARCS-De-Identified/npsr-cm47/data>
+
+**From website:**
+
+The Statewide Planning and Research Cooperative System (SPARCS) Inpatient De-identified File contains discharge level detail on patient characteristics, diagnoses, treatments, services, and charges.
+This data file contains basic record level detail for the discharge.
+The de-identified data file does not contain data that is protected health information (PHI) under HIPAA.
+The health information is not individually identifiable; all data elements considered identifiable have been redacted.
+For example, the direct identifiers regarding a date have the day and month portion of the date removed.
+
+Discharges that are identified as Abortion will have some information redacted; Facility ID, Facility Name, Health Service Area, Hospital County, Operating Certificate, and all provider license numbers.
+Patient zip code is limited to the first three digits.
+Zip codes for population size less than 20,000 are blank.
+Out of State are "OOS'.
+Patient age is presented in age groups: 0 to 17, 18 to 29, 30 to 49, 50 to 69, and 70 or Older.
+
+**Logistics:**
+
+-   2,428,905 rows of patient data
+
+-   34 columns
+
+    -   Health Service Area
+    -   Hospital County
+    -   Operating Certificate Number
+    -   Facility ID, Facility Name
+    -   Age Group, Zip Code (3 digit)
+    -   Gender
+    -   Race
+    -   Ethnicity
+    -   Length of Stay
+    -   Type of Admission
+    -   Patient Disposition
+    -   Discharge Year
+    -   CCS Diagnosis Code
+    -   CCS Diagnosis Description
+    -   CCS Procedure Code
+    -   CCS Procedure Description
+    -   APR DRG Code
+    -   APR DRG Description
+    -   APR MDC Code
+    -   APR MDC Description
+    -   APR Severity of Illness Code
+    -   APR Severity of Illness
+    -   APR Risk of Mortality
+    -   APR Medical Surgical Description
+    -   Payment Typology
+    -   Payment Typology
+    -   Birth Weight,
+    -   Abortion Edit Indicator
+    -   Emergency Department
+    -   Total Charges
+    -   Total Costs
+
+## Visualizations and Plots
+
+#### **Plot 1: Price vs Length of Stay**
 

### 2023-09-23 STAT405Project.qmd
fixed some other stuff
added statistical testing
updated pie plot

@@
 
-*(last updated 09/21/2023)*
+*(last updated 09/23/2023)*
 
@@
 lmfs <- lm(y ~ length_of_stay)
+summary(lmfs)
 abline(lmfs, col = "RED")
@@
 pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA)
-legend("left", legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
+legend("topleft", legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
 ```

### 2023-09-23 STAT405Project.qmd
Update STAT405Project.qmd

@@
 ---
-##
 
@@
 
-## Visualizations and Plots
+# Visualizations and Plots
 
-#### **Plot 1: Price vs Length of Stay**
+## **Plot 1: Price vs Length of Stay and Linear Regression Statistical Test**
 
@@
 ```{r}
+#| echo: false
 #install.packages("readr")
@@
 ```{r}
+#| echo: false
 y <- cleaned_data$`Total Charges`
@@
 ```{r}
+#| echo: false
 length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
@@
 ```
+### Linear Regression Statistical Test and Analysis
+
+In our analysis of patient data for NYC in 2013, we conducted a linear regression to explore the impact of the length of stay on the associated price.
+The results revealed a statistically significant and strong positive relationship between these variables.
+The coefficient for 'length_of_stay' was estimated at 6140.73 with a standard error of 6.622 and a remarkably high t-value of 927.39.
+This suggests that as the length of stay increases, there is a substantial increase in the price, a relationship supported by a p-value close to zero.
+The R-squared value of 0.5567 indicates that approximately 55.67% of the price variation can be explained by the length of stay, demonstrating its relevance as a predictor.
+These findings provide important insights for healthcare management and policymaking, highlighting the need to consider and manage the cost implications of varying lengths of stay for NYC patients in 2013.
+
+### Plot Desciption and Inferencing
+
+In addition to the regression analysis, we visualized the relationship between the length of stay and the associated price using a scatter plot with a fitted regression line.
+The scatter plot provides a clear graphical representation of the positive correlation between these variables.
+As the length of stay increases, there is a discernible upward trend in price.
+The regression line, which is based on the linear regression model we performed, further accentuates this trend.
+It serves as a useful tool for illustrating the direction and strength of the relationship, making it easier to interpret the practical implications of our findings.
+This visualization reinforces the statistical significance we observed in our analysis and enhances our understanding of the relationship between patient length of stay and healthcare costs in NYC in 2013.
+
+## **Plot 2: Boxplot of Total Costs**
 
 ```{r}
+#| echo: false
+cleaned_data$'Total Costs'[cleaned_data$'Total Costs' > quantile(cleaned_data$'Total Costs', 0.95)] <- quantile(cleaned_data$'Total Costs', 0.95)
 
 boxplot(cleaned_data$'Total Costs', 
-        main="Box Plot of Total Costs",
-        ylab= "Total costs")
+        main="Box Plot of Total Costs (Outliers Truncated)",
+        ylab="Total Costs")
 ```
+### Plot Desciption and Inferencing
 
-```{r}
+## **Plot 3: Payment Type Distribution**
+```{r warning = FALSE}
+#| echo: false
 options(repr.plot.width = 100, repr.plot.height = 100)
@@
 
+## **Plot 4: Trends in Total Charges and Length of Stay**
 ```{r}
+#| echo: false
 
-# Create the line chart
 plot(cleaned_data$`Length of Stay`, cleaned_data$`Total Charges`, type = "l", col = "blue", xlab = "Length of Stay", ylab = "Total Charges", main = "Trends in Total Charges and Length of Stay")
 
-# Add a line for Total Costs
 lines(cleaned_data$`Birth Weight`, cleaned_data$`Total Costs`, col = "red")
 
-# Add a legend
 legend("topright", legend = c("Total Charges", "Total Costs"), col = c("blue", "red"), lty = 1)
 
-# Add a grid
 grid()
@@
 ```
+
+## **Plot 5: Heatmap of Gender vs. Age Group**
+
 ```{r}
-
+#| echo: false
 heatmap_data <- table(cleaned_data$`Age Group`, cleaned_data$Gender)
@@
 ```
+
+## **Plot 6: Barplot of Age Group Distribution of Patients**
 ```{r}
+#| echo: false
 ageGroups <- table(cleaned_data$`Age Group`)

### 2023-09-23 STAT405Project.qmd
horizontal boxplot

@@
         main="Box Plot of Total Costs (Outliers Truncated)",
-        ylab="Total Costs")
+        ylab="Total Costs",
+        horizontal = TRUE)
 ```
 ### Plot Desciption and Inferencing 
+The following boxplot refers to our Total Cost data and is representative of our data that has been cleaned to remove outliers that are in our top 5% of data points, meaning that the outliers that are extremely high are cleaned out.
 
@@
 ## **Plot 4: Trends in Total Charges and Length of Stay**
-```{r}
+```{r warning = FALSE}
 #| echo: false

### 2023-09-23 STAT405Project.qmd
all writeups

@@
 
-In our analysis of patient data for NYC in 2013, we conducted a linear regression to explore the impact of the length of stay on the associated price.
+In our analysis of patient data for the state of New York in 2013, we conducted a linear regression to explore the impact of the length of stay on the associated price.
 The results revealed a statistically significant and strong positive relationship between these variables.
@@
 The R-squared value of 0.5567 indicates that approximately 55.67% of the price variation can be explained by the length of stay, demonstrating its relevance as a predictor.
-These findings provide important insights for healthcare management and policymaking, highlighting the need to consider and manage the cost implications of varying lengths of stay for NYC patients in 2013.
+These findings provide important insights for healthcare management and policymaking, highlighting the need to consider and manage the cost implications of varying lengths of stay for New York state patients in 2013.
 
@@
 It serves as a useful tool for illustrating the direction and strength of the relationship, making it easier to interpret the practical implications of our findings.
-This visualization reinforces the statistical significance we observed in our analysis and enhances our understanding of the relationship between patient length of stay and healthcare costs in NYC in 2013.
+This visualization reinforces the statistical significance we observed in our analysis and enhances our understanding of the relationship between patient length of stay and healthcare costs for patients in the state of New York in 2013.
 
@@
 cleaned_data$'Total Costs'[cleaned_data$'Total Costs' > quantile(cleaned_data$'Total Costs', 0.95)] <- quantile(cleaned_data$'Total Costs', 0.95)
+
+summary(cleaned_data$'Total Costs')
 
@@
 The following boxplot refers to our Total Cost data and is representative of our data that has been cleaned to remove outliers that are in our top 5% of data points, meaning that the outliers that are extremely high are cleaned out.
+In our analysis, we used a horizontal box plot to summarize the distribution of 'Total Costs' in our dataset.
+The median, located at approximately $8,247.96, divides the data into two halves, with the interquartile range (IQR) spanning from Q1 at $4,519.54 to Q3 at $15,840.81.
+Outliers, representing data points significantly deviating from the central distribution, are visible.
+The total cost range extends from $0.41 to $45,292.36, showcasing the data's spread.
+The mean total cost is around $12,360.20, but outliers can influence this value.
+The boxplot also is skewed to the right.
+This visualization provides key insights into our cost structure analysis.
+
+From this we can infer that there are many outliers, however, in the discussion of costs they are heavily spread and that most patients are paying within our IQR but there are many different outliers that spread when we are discussing costs for healthcare.
 
@@
 payments <- table(cleaned_data$`Payment Typology 1`)
-pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA)
+pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA, )
 legend("topleft", legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
+
 ```
+### Plot Desciption and Inferencing 
+The following piechart identifies the payment type distribution for patients in New York in 2013.
+As we can observe, most patients are using Medicare as a way to pay for their medical bills followed by Medicaid.
+These government programs are important to New York residents and provides important data on the impact of healthcare programs.
+The exact percentagse are Medicaid at 59.7% and Medicaid at 26.7%, followed by Private Insurance at around 10%.
+This information reveals how important the impact of governmental insurance programs are to individuals in New York.
 
@@
 grid()
+```
+### Plot Desciption and Inferencing 
+The plotted data and corresponding projections unmistakably depict a consistent total cost profile, irrespective of variations in both the length of stay and total charges.
+This apparent lack of correlation between the length of stay and associated costs prompts a deeper inquiry into the potential contributing factors.
+It would be important to explore the influence of fixed operational costs, negotiated contractual agreements, and astute cost management strategies that may explain this observed financial consistancy between the costs for a hospital and for its patients
 
-#Total cost doesn't change even if length of stay increases and total charge increases.
+## **Plot 5: Barplot of Age Group Distribution of Patients**
+```{r}
+#| echo: false
+ageGroups <- table(cleaned_data$`Age Group`)
+barplot(ageGroups, main = "Age Group Distribution", xlab = "Age Group", 
+        ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)) 
+```
+### Plot Desciption and Inferencing 
+Another important piece of data that we wanted to observe is the amount of patients from each age group of the population.
+The data is divided into five segmests 0-17, 18-29, 30-49, 50-69, and 70-older.
+As we can see there is an upward trend of the amount of patients as the ages of the patients increases, although there is a small decrease between 0-17 and 8-29.
+The largest frequency if the 70-older age bucket which has around 289248 compared to 18-29 with our smallest frequency of 53009.This increase in older patients reveals the impact of older populations in New York and the necesity to increase health care efforts towards older populations.
 
-```
-
-## **Plot 5: Heatmap of Gender vs. Age Group**
+## **Plot 6: Heatmap of Gender vs. Age Group**
 
@@
 ```
-
-## **Plot 6: Barplot of Age Group Distribution of Patients**
-```{r}
-#| echo: false
-ageGroups <- table(cleaned_data$`Age Group`)
-barplot(ageGroups, main = "Age Group Distribution", xlab = "Age Group", 
-        ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)) 
-```
+### Plot Desciption and Inferencing 
+The following heatmap shows the relationship between the gender and age group for patients in New York in 2013.
+As we can observe, women tend to be patients at hospitals at a higher rate then men.
+Especially for the 70 or older bracket which is the highest frequency of hospital patients, females are significantly a larger proprotion of who are patients compared to men.
+Another important data point to understand is the identification of "U" or individuals who do not identify as male and female, which is high for ages 18-29, which is an important data point to help increase healthcare for individuals with fluid gender idenity or may identify as non-binary.
 

### 2023-09-26 STAT405Project.qmd
9/26 finished

@@
 ---
-
 *(last updated 09/23/2023)*
@@
 ```
+
 ### Linear Regression Statistical Test and Analysis
@@
 ```
+
 ### Plot Desciption and Inferencing
+
 The following boxplot refers to our Total Cost data and is representative of our data that has been cleaned to remove outliers that are in our top 5% of data points, meaning that the outliers that are extremely high are cleaned out.
 In our analysis, we used a horizontal box plot to summarize the distribution of 'Total Costs' in our dataset.
-The median, located at approximately $8,247.96, divides the data into two halves, with the interquartile range (IQR) spanning from Q1 at $4,519.54 to Q3 at $15,840.81.
+The median, located at approximately \$8,247.96, divides the data into two halves, with the interquartile range (IQR) spanning from Q1 at \$4,519.54 to Q3 at \$15,840.81.
 Outliers, representing data points significantly deviating from the central distribution, are visible.
-The total cost range extends from $0.41 to $45,292.36, showcasing the data's spread.
-The mean total cost is around $12,360.20, but outliers can influence this value.
+The total cost range extends from \$0.41 to \$45,292.36, showcasing the data's spread.
+The mean total cost is around \$12,360.20, but outliers can influence this value.
 The boxplot also is skewed to the right.
@@
 ## **Plot 3: Payment Type Distribution**
-```{r warning = FALSE}
+
+```{r, warning = FALSE, fig.height = 6, fig.width = 50}
 #| echo: false
@@
 pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA, )
-legend("topleft", legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
+legend(1,1,1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
 
 ```
+
 ### Plot Desciption and Inferencing
+
 The following piechart identifies the payment type distribution for patients in New York in 2013.
@@
 ## **Plot 4: Trends in Total Charges and Length of Stay**
+
 ```{r warning = FALSE}
@@
 ```
+
 ### Plot Desciption and Inferencing
+
 The plotted data and corresponding projections unmistakably depict a consistent total cost profile, irrespective of variations in both the length of stay and total charges.
@@
 ## **Plot 5: Barplot of Age Group Distribution of Patients**
+
 ```{r}
@@
 ```
+
 ### Plot Desciption and Inferencing
+
 Another important piece of data that we wanted to observe is the amount of patients from each age group of the population.
@@
 ```
+
 ### Plot Desciption and Inferencing
+
 The following heatmap shows the relationship between the gender and age group for patients in New York in 2013.
@@
 Especially for the 70 or older bracket which is the highest frequency of hospital patients, females are significantly a larger proprotion of who are patients compared to men.
-Another important data point to understand is the identification of "U" or individuals who do not identify as male and female, which is high for ages 18-29, which is an important data point to help increase healthcare for individuals with fluid gender idenity or may identify as non-binary.
+Another important data point to understand is the identification of "U" or individuals who do not identify as male and female, which is high for ages 18-29, which is an important data point to help increase healthcare for individuals with fluid gender identity or may identify as non-binary.
 
 <br />
+
+## Plot 7: Stacked Bar Chart by Gender and Race
+
+```{r}
+#| echo: false
+#install.packages("ggplot2")
+library(ggplot2)
+ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
+  geom_bar() +
+  labs(
+    title = "Stacked Bar Chart by Gender and Race",
+    x = "Gender",
+    y = "Count"
+  ) +
+  scale_fill_manual(
+    values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
+    name = "Race"
+  ) +
+  scale_y_continuous(labels = scales::comma)
+  theme_minimal()
+
+```
+
+## **Plot 8:
+```{r}
+#| echo: false
+#install.packages("ggplot2")
+library(ggplot2)
+ggplot(cleaned_data) + 
+        aes("Race", "Total Costs", color = "salmon") + 
+        geom_jitter()
+
+```

### 2023-09-28 STAT405Project.qmd
Update STAT405Project.qmd

@@
 ---
+
 *(last updated 09/23/2023)*
@@
 #install.packages("readr")
+#install.packages("xtable")
 library(readr)
-
+library(xtable)
+options(xtable.floating = FALSE)
+options(xtable.timestamp = "")
 data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)
@@
 
-```{r}
+```{r, results='asis'}
 #| echo: false
@@
 lmfs <- lm(y ~ length_of_stay)
-summary(lmfs)
+xtable(lmfs, comment = FALSE)
 abline(lmfs, col = "RED")
@@
   ) +
-  scale_y_continuous(labels = scales::comma)
   theme_minimal()
@@
 
-## **Plot 8:
-```{r}
+## Plot 8: Boxplots on the Total Costs on the Risk of Mortality
+
+```{r warning = FALSE}
 #| echo: false
 #install.packages("ggplot2")
-library(ggplot2)
 ggplot(cleaned_data) + 
-        aes("Race", "Total Costs", color = "salmon") + 
-        geom_jitter()
+  aes(x = cleaned_data$'APR Risk of Mortality', y = cleaned_data$"Total Costs", color = "salmon") + 
+  geom_boxplot() +
+  labs(title = "Total Costs on Risk of Mortality",
+       x = "APR Risk of Mortality",  # X-axis title
+       y = "Total Costs")  # Y-axis title
 
 ```
+
+## Plot 9: Boxplots on the Total Costs on the Risk of Mortality
+
+```{r warning = FALSE}
+#| echo: false
+
+```
+
+## Plot 10:
+
+```{r warning = FALSE}
+#| echo: false
+
+```
+
+## Plot 11:
+
+```{r warning = FALSE}
+#| echo: false
+
+```
+
+## Plot 12:
+
+```{r warning = FALSE}
+#| echo: false
+
+```

### 2023-09-30 STAT405Project.qmd
Added plots

@@
 
-## Plot 9: Boxplots on the Total Costs on the Risk of Mortality
+## Plot 9:
 
@@
 #| echo: false
+ggplot(cleaned_data, aes(x = cleaned_data$`Length of Stay`, y = cleaned_data$`Total Charges`))+
+  geom_point(size = 0.2) +
+  facet_grid(Race ~ `Type of Admission`) +
+  labs(title = "Length of Stay v. Total Charges", x = "Length of Stay", y = "Total Charges")
 
@@
 #| echo: false
+ggplot(cleaned_data, aes(x = `Total Costs`)) +
+  geom_density(fill = "blue", alpha = 0.5) +
+  facet_grid(`Age Group` ~ .)
+  labs(title = "Density Plot Example", x = "X-axis Label", y = "Density")
 

### 2023-09-30 STAT405Project.qmd
added plot

@@
 #| echo: false
-#install.packages("readr")
-#install.packages("xtable")
+# install.packages("readr")
+# install.packages("xtable")
 library(readr)
@@
 options(xtable.timestamp = "")
-data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)
+data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv", show_col_types = FALSE)
 
@@
 #| echo: false
-cleaned_data$'Total Costs'[cleaned_data$'Total Costs' > quantile(cleaned_data$'Total Costs', 0.95)] <- quantile(cleaned_data$'Total Costs', 0.95)
+cleaned_data$"Total Costs"[cleaned_data$"Total Costs" > quantile(cleaned_data$"Total Costs", 0.95)] <- quantile(cleaned_data$"Total Costs", 0.95)
 
-summary(cleaned_data$'Total Costs')
+summary(cleaned_data$"Total Costs")
 
-boxplot(cleaned_data$'Total Costs', 
-        main="Box Plot of Total Costs (Outliers Truncated)",
-        ylab="Total Costs",
-        horizontal = TRUE)
+boxplot(cleaned_data$"Total Costs",
+    main = "Box Plot of Total Costs (Outliers Truncated)",
+    ylab = "Total Costs",
+    horizontal = TRUE
+)
 ```
@@
 payments <- table(cleaned_data$`Payment Typology 1`)
-pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA, )
-legend(1,1,1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
+pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA, )
+legend(1, 1, 1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
 
@@
 ageGroups <- table(cleaned_data$`Age Group`)
-barplot(ageGroups, main = "Age Group Distribution", xlab = "Age Group", 
-        ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)) 
+barplot(ageGroups,
+    main = "Age Group Distribution", xlab = "Age Group",
+    ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)
+)
 ```
@@
     labRow = row_names,
-        labCol = col_names)
+    labCol = col_names
+)
 
@@
 #| echo: false
-#install.packages("ggplot2")
+# install.packages("ggplot2")
 library(ggplot2)
@@
 #| echo: false
-#install.packages("ggplot2")
+# install.packages("ggplot2")
 ggplot(cleaned_data) +
-  aes(x = cleaned_data$'APR Risk of Mortality', y = cleaned_data$"Total Costs", color = "salmon") + 
+    aes(x = cleaned_data$"APR Risk of Mortality", y = cleaned_data$"Total Costs", color = "salmon") +
     geom_boxplot() +
-  labs(title = "Total Costs on Risk of Mortality",
+    labs(
+        title = "Total Costs on Risk of Mortality",
         x = "APR Risk of Mortality", # X-axis title
-       y = "Total Costs")  # Y-axis title
+        y = "Total Costs"
+    ) # Y-axis title
 
@@
 #| echo: false
+install.packages("ggExtra")
+
+set.seed(100)
+library(ggplot2)
+library(ggExtra)
+library(dplyr)
+random_samples <- sample_n(cleaned_data, 100, replace = FALSE)
+
+Costs <- random_samples$`Total Costs`
+Charges <- random_samples$`Total Charges`
+
+g <- ggplot(random_samples, aes(Costs, Charges, howdy)) +
+    geom_count(size = 1) +
+    geom_smooth(method = "lm", se = F, color = "purple") +
+    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
+    ggtitle("Total Costs vs Total Charges for 100 Patients")
+
+ggMarginal(g, type = "histogram")
 
@@
 #| echo: false
-ggplot(cleaned_data, aes(x = cleaned_data$`Length of Stay`, y = cleaned_data$`Total Charges`))+
+ggplot(cleaned_data, aes(x = cleaned_data$`Length of Stay`, y = cleaned_data$`Total Charges`)) +
     geom_point(size = 0.2) +

### 2023-10-01 STAT405Project.qmd
basically a lot of shit

--- 2023-09-28 STAT405Project.qmd
@@
 
-*(last updated 09/23/2023)*
+*(last updated 09/30/2023)*
 
@@
 
+<br />
+
+\newpage
+
 # Visualizations and Plots
@@
 
-```{r}
+```{r, include = FALSE}
 #| echo: false
-#install.packages("readr")
-#install.packages("xtable")
+# install.packages("readr")
+# install.packages("xtable")
 library(readr)
+library(dplyr)
 library(xtable)
@@
 length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
+
+# Set scipen to a large value to avoid scientific notation
+options(scipen = 10)
+
+# Create a scatter plot
 plot(length_of_stay, y, xlab = "Length of Stay", main = "Price vs Length of Stay", ylab = "Price")
 
+# Reset scipen to its default value after the plot
+options(scipen = 0)
+
 lmfs <- lm(y ~ length_of_stay)
-xtable(lmfs, comment = FALSE)
 abline(lmfs, col = "RED")
+
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 1a: Length of Stay on Price Regression Dotplot</span>
+
+```{r, results='asis'}
+#| echo: false
+table_lmfs <- xtable::xtable(lmfs, comment = NA, caption = " Figure 1.b: Linear Model Summary")
+
+cat("\\begin{center}\n")
+print(table_lmfs, comment = FALSE)
+cat("\\end{center}\n")
+```
+<i>Figure 1b: Table of Linear Regression Modeling Data</i>
 
@@
 
+\newpage
+
 ## **Plot 2: Boxplot of Total Costs**
@@
 #| echo: false
-cleaned_data$'Total Costs'[cleaned_data$'Total Costs' > quantile(cleaned_data$'Total Costs', 0.95)] <- quantile(cleaned_data$'Total Costs', 0.95)
+cleaned_data$"Total Costs"[cleaned_data$"Total Costs" > quantile(cleaned_data$"Total Costs", 0.95)] <- quantile(cleaned_data$"Total Costs", 0.95)
 
-summary(cleaned_data$'Total Costs')
+#summary(cleaned_data$"Total Costs")
 
-boxplot(cleaned_data$'Total Costs', 
-        main="Box Plot of Total Costs (Outliers Truncated)",
-        ylab="Total Costs",
-        horizontal = TRUE)
+boxplot(cleaned_data$"Total Costs",
+    main = "Box Plot of Total Costs (Outliers Truncated)",
+    ylab = "Total Costs",
+    horizontal = TRUE
+)
+
 ```
+<i>Figure 2: Total Costs Boxplot</i>
 
@@
 
+\newpage
+
 ## **Plot 3: Payment Type Distribution**
@@
 #| echo: false
-options(repr.plot.width = 100, repr.plot.height = 100)
+# Set the margin to include more space at the bottom
+par(mar = c(5, 4, 4, 2) + 0.1)
 
@@
 payments <- table(cleaned_data$`Payment Typology 1`)
-pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA, )
-legend(1,1,1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
+pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA)
+legend(1, 1, 1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
 
 ```
+<i>Figure 3: Pie Chart of Payment Distribution</i>
 
@@
 This information reveals how important the impact of governmental insurance programs are to individuals in New York.
+
+\newpage
 
@@
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 4: Trends of Total Charges and Legnth of Stay </span>
 
@@
 
+\newpage
+
 ## **Plot 5: Barplot of Age Group Distribution of Patients**
@@
 ageGroups <- table(cleaned_data$`Age Group`)
-barplot(ageGroups, main = "Age Group Distribution", xlab = "Age Group", 
-        ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)) 
+barplot(ageGroups,
+    main = "Age Group Distribution", xlab = "Age Group",
+    ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)
+)
+mtext("Figure 5: Age Group and Distribtion", side = 1, line = 4, adj = -0.1, cex = 0.7, font = 3, col = "black")
+
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 5: Age Group Distribution of Patients in New York</span>
 
@@
 The largest frequency if the 70-older age bucket which has around 289248 compared to 18-29 with our smallest frequency of 53009.This increase in older patients reveals the impact of older populations in New York and the necesity to increase health care efforts towards older populations.
+
+\newpage
 
@@
     labRow = row_names,
-        labCol = col_names)
+    labCol = col_names
+)
+mtext("Figure 6: Gender vs. Age Group Heatmap", side = 1, line = 4, adj = -0.1, cex = 0.7, font = 3, col = "black")
 
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 6: Gender vs Age Group Heatmapt</span>
 
@@
 
-<br />
+\newpage
 
@@
 #| echo: false
-#install.packages("ggplot2")
+# install.packages("ggplot2")
 library(ggplot2)
-ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
+
+plot7 <- ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
   geom_bar() +
@@
   scale_fill_manual(
-    values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
+    values = c("White" = "lightblue", "African American" = "lightcoral", "Other" = "lightgreen"),
+    breaks = c("White", "Black", "Other"),
+    labels = c("White", "Black", "Other"),
     name = "Race"
@@
 
+plot7 + ggtitle("Data Source: mtcars dataset, ggplot2 package")
+
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 7: Stacked Bar Chart by Race and Gender</span>
+
+## Plot Description and Inferencing
+
+In the analysis of patient demographics in the year 2013, a bar chart was employed to visually represent the breakdown of patients by gender and further stratify each gender by race.
+A notable observation from the chart is the significant predominance of female patients over male patients during this period.
+
+Upon closer examination of the male demographic, the distribution reveals a relatively balanced representation between individuals identified as white and those classified as not white.
+This suggests a relatively even distribution among male patients concerning racial categories.
+In contrast, the female demographic exhibits a slightly more diverse distribution across racial categories.
+This observation implies that, among female patients, there is a greater degree of heterogeneity in terms of racial composition compared to their male counterparts.
+
+These findings contribute valuable insights into the gender and racial composition of the patient population in 2013, emphasizing the need for a nuanced understanding of demographic trends within healthcare settings.
+Further exploration and analysis of these demographic patterns may reveal additional insights relevant to healthcare planning, resource allocation, and the provision of patient-centered care.
+
+\newpage
 
@@
 #| echo: false
-#install.packages("ggplot2")
+# install.packages("ggplot2")
 ggplot(cleaned_data) +
-  aes(x = cleaned_data$'APR Risk of Mortality', y = cleaned_data$"Total Costs", color = "salmon") + 
+    aes(x = cleaned_data$"APR Risk of Mortality", y = cleaned_data$"Total Costs", color = "salmon") +
     geom_boxplot() +
-  labs(title = "Total Costs on Risk of Mortality",
+    labs(
+        title = "Total Costs on Risk of Mortality",
         x = "APR Risk of Mortality", # X-axis title
-       y = "Total Costs")  # Y-axis title
+        y = "Total Costs"
+    ) # Y-axis title
 
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 8: Boxplots on Total Cost and Mortality</span>
 
-## Plot 9: Boxplots on the Total Costs on the Risk of Mortality
+## Plot Description and Inferencing
+In Figure 8, our objective was to assess the central tendency of total hospital costs across different levels of Risk of Mortality.
+Employing separate box plots for each severity category allowed us to discern distinct patterns in the distribution of total costs associated with varying degrees of mortality risk.
+Notably, when focusing on Extreme cases of mortality, a clear trend emerged—median total costs were significantly higher, accompanied by a broader spread in the distribution.
+This observation implies that patients facing an Extreme risk of mortality not only incur higher median costs but also exhibit greater variability in the financial implications of their healthcare, suggesting a more diverse range of medical interventions and resource utilization.
+
+This finding aligns with our initial expectation that higher levels of risk of mortality correspond to elevated costs of care.
+The discernible trend in the box plots reinforces the intuitive understanding that patients with a greater severity of illness often require more extensive and resource-intensive medical interventions, leading to increased healthcare costs.
+Beyond its immediate implications, this nuanced analysis offers valuable insights for healthcare planning and resource allocation, emphasizing the need for tailored strategies based on the specific needs associated with varying risk levels.
+This approach contributes to more effective and efficient healthcare delivery, promoting a patient-centered and economically sustainable healthcare system.
+
+\newpage
+
+## Plot 9: Violin Plot Comparing New York City and Western New York's Severity of Illness
 
@@
 #| echo: false
+subsetted_data <- cleaned_data[cleaned_data$`Health Service Area` %in% c("New York City", "Western NY"), ]
+
+# Load the ggplot2 library
+library(ggplot2)
+
+# Create a violin plot for APR Severity of Illness Code
+ggplot(subsetted_data, aes(x = factor(`Health Service Area`), y = `APR Severity of Illness Code`, fill = `Health Service Area`)) +
+  geom_violin() +
+  labs(
+    x = "Health Service Area",
+    y = "APR Severity of Illness Code",
+    fill = "Health Service Area",
+    title = "Distribution of APR Severity of Illness Code in New York City"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 9: NYC and Western NY Severity of Illness Comparsion </span>
 
-## Plot 10:
+## Plot Description and Inferencing
+Figure 9 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
+The focus is on the distribution of patients' APR severity codes, measured on a 1 to 4 scale (1 being mild, 2 being moderate, 3 being serious, and 4 being extreme).
+Notably, the violin plot reveals a higher frequency of APR severity code 2 assignments for patients from NYC compared to WNY.
+This divergence prompts exploration into potential variations in labeling practices between the two regions.
+However, beyond this distinction, the overall shapes of the violin distributions appear similar, suggesting a general alignment in severity profiles.
+This observation implies a degree of consistency in patient conditions based on APR severity codes, despite the regional differences in labeling practices.
+Understanding such variations is crucial for refining medical coding precision and ensuring uniformity in patient severity assessments across diverse geographical contexts.
+
+\newpage
+
+## Plot 10: Distribution of APR Severity of Illness for Medicaid Patients
+
+```{r, warning = FALSE}
+#| echo: false
+# Load the required libraries
+library(ggplot2)
+
+medicaid_data <- cleaned_data %>%
+  filter(`Payment Typology 1` == "Medicaid")
+
+# Create a ridgeline plot to visualize APR Severity of Illness
+ggplot(medicaid_data, aes(x = `APR Severity of Illness Code`, y = ..density.., fill = `APR Severity of Illness Code`)) +
+  geom_density(alpha = 0.6) +
+  labs(
+    x = "APR Severity of Illness Code",
+    y = "Density",
+    fill = "APR Severity of Illness Code",
+    title = "Distribution of APR Severity of Illness for Medicaid Patients"
+  ) +
+  theme_minimal() +
+  scale_fill_brewer(palette = "Set1")
+
+```
+<span style="font-size: 14px; font-style: italic;">Figure 10: APR Severity of Illness for Medicaid Patients </span>
+
+## Plot Description and Inferencing
+In our examination of Figure 10, our focus shifted toward a specific subset of patients—those under Medicaid care.
+Medicaid, a health program designed to support individuals classified as economically disadvantaged, prompted our investigation into the severity of diagnoses within this demographic.
+Our objective was to discern whether patients under Medicaid receive comparable attention and care in contrast to their counterparts outside of the program.
+
+Distinct from the broader population, our analysis revealed a notable prevalence of patients with "mild" severity codes among those under Medicaid care.
+This intriguing observation necessitates careful consideration of various contributing factors.
+While we refrain from making definitive conclusions about the intentional prioritization of severity levels for Medicaid patients, this finding presents a compelling avenue for further exploration.
+
+The presence of a higher number of "mild" severity cases within the Medicaid subset raises questions about potential disparities in healthcare provision.
+It is important to acknowledge that a comprehensive understanding of this phenomenon requires additional contextual data.
+Future investigations could delve deeper into the intricate dynamics surrounding patient severity levels, ensuring a nuanced exploration that considers various influencing factors.
+
+\newpage
+
+## Plot 11: Facet Grid of Dotplots for Total Charges on Length of Stay
 
@@
 #| echo: false
+library(ggplot2)
+
+ggplot(cleaned_data, aes(x = cleaned_data$`Length of Stay`, y = cleaned_data$`Total Charges`)) +
+  geom_point(size = 0.2) +
+  facet_grid(Race ~ `Type of Admission`) +
+  labs(title = "Length of Stay v. Total Charges", x = "Length of Stay", y = "Total Charges") +
+  scale_y_continuous(labels = scales::comma, 
+                     breaks = scales::pretty_breaks(), 
+                     limits = c(0, max(cleaned_data$`Total Charges`)), # 
+                     expand = c(0, 0)) +
+  theme(axis.text = element_text(size = 8))
 
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 11: Dotplots for Total Charges on Length of Stay </span>
 
-## Plot 11:
+## Plot Description and Inferencing
+
+In Figure 11, our aim was to dissect assessments based on both race and the type of admission.
+Through our analysis, a discernible pattern emerged, revealing consistent spikes in each category of total charges, particularly towards the lower end of the length of stay spectrum.
+This observation suggests that, irrespective of race, there are commonalities in the distribution of total charges, with pronounced concentration points occurring for shorter stays.
+
+Upon closer examination, the data points appear to be uniformly distributed across all racial categories, implying a degree of similarity in the cost distribution patterns among different racial groups.
+Interestingly, the primary differentiating factor contributing to variations in total costs is identified as the type of admission.
+This finding underscores the significance of admission characteristics in influencing healthcare costs, emphasizing that differences in total charges are more prominently linked to variations in admission types rather than racial distinctions.
+
+These insights shed light on the nuanced interplay between race, type of admission, and healthcare costs.
+Understanding these relationships is crucial for healthcare administrators and policymakers in designing targeted strategies for cost-effective care delivery, ensuring equitable access, and fostering a healthcare system that is both inclusive and economically sustainable.
+
+\newpage
+
+## Plot 12: Facet Density Plots of Total Cost per Age-Group Buckets
 
@@
 #| echo: false
+ggplot(cleaned_data, aes(x = `Total Costs`)) +
+  geom_density(fill = "blue", alpha = 0.5) +
+  facet_grid(`Age Group` ~ .) +
+labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density");
 
 ```
+<span style="font-size: 14px; font-style: italic;">Figure 12: Density Plots of Total Costs per Age-Groups</span>
 
-## Plot 12:
+## Plot Description and Inferencing
+In Figure 12, we analyzed age-specific total costs by charting density plots using a facet grid.
+For patients under 17, costs skewed right, suggesting lower expenses, potentially due to the overall good health of children.
+In contrast, as age increases, a broader and more dispersed peak in total costs emerges, indicating that older patients require more extensive and potentially costlier healthcare interventions, such as hospice services.
+This exploration sheds light on age-related patterns in healthcare costs, guiding strategies for resource allocation and care planning across different life stages.
 
-```{r warning = FALSE}
-#| echo: false
+\newpage
 
-```
+### References
+New York State Department of Health.
+"Hospital Inpatient Discharges (SPARCS De-Identified)." data.ny.gov, 2013, https://health.data.ny.gov/Health/Hospital-Inpatient-Discharges-SPARCS-De-Identified/npsr-cm47.

### 2023-10-01 STAT405Project.qmd
YAY

@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 4: Trends of Total Charges and Legnth of Stay </span>
+<span style="font-size: 14px; font-style: italic;">Figure 4:Trends of Total Charges and Legnth of Stay </span>
 
@@
 )
-mtext("Figure 5: Age Group and Distribtion", side = 1, line = 4, adj = -0.1, cex = 0.7, font = 3, col = "black")
 
@@
 )
-mtext("Figure 6: Gender vs. Age Group Heatmap", side = 1, line = 4, adj = -0.1, cex = 0.7, font = 3, col = "black")
 
@@
 ggplot(cleaned_data) +
-    aes(x = cleaned_data$"APR Risk of Mortality", y = cleaned_data$"Total Costs", color = "salmon") +
-    geom_boxplot() +
+  aes(x = `APR Risk of Mortality`, y = `Total Costs`) +
+  geom_boxplot(color = "salmon") +
   labs(
     title = "Total Costs on Risk of Mortality",
-        x = "APR Risk of Mortality", # X-axis title
+    x = "APR Risk of Mortality",
     y = "Total Costs"
-    ) # Y-axis title
+  ) +
+  theme_minimal()
 
@@
 #| echo: false
-ggplot(cleaned_data, aes(x = `Total Costs`)) +
-  geom_density(fill = "blue", alpha = 0.5) +
+gg <- ggplot(cleaned_data, aes(x = `Total Costs`, y = ..density.., fill = `Age Group`)) +
+  geom_density(alpha = 0.5) +
   facet_grid(`Age Group` ~ .) +
-labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density");
+  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density")
 
+gg + scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
+  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density (in thousands)") +
+  theme_minimal()
 ```

### 2023-10-08 NewSTAT405.qmd
Update NewSTAT405.qmd

--- 2023-10-01 STAT405Project.qmd
@@
 
-*(last updated 09/30/2023)*
+*(last updated 10/07/2023)*
 
@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 1a: Length of Stay on Price Regression Dotplot</span>
+<p align = "center">
+[Figure 1a: Length of Stay on Price Regression Dotplot]
+</p>
 
@@
 ```
-<i>Figure 1b: Table of Linear Regression Modeling Data</i>
+<p align = "center">
+[Figure 1b: Table of Linear Regression Modeling Data]
+</p>
 
@@
 ```
-<i>Figure 2: Total Costs Boxplot</i>
+<p align = "center">
+[Figure 2: Total Costs Boxplot]
+</p>
 
@@
 
-```{r, warning = FALSE, fig.height = 6, fig.width = 50}
+```{r, warning = FALSE, fig.height = 15, fig.width = 50}
 #| echo: false
@@
 par(mar = c(5, 4, 4, 2) + 0.1)
-
+par(cex.main = 5 )
 cleaned_data$`Payment Typology 1` <- as.factor(cleaned_data$`Payment Typology 1`)
@@
 pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA)
-legend(1, 1, 1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)
+legend(1, 1, 1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 3)
 
 ```
-<i>Figure 3: Pie Chart of Payment Distribution</i>
+<p align = "center">
+[Figure 3: Pie Chart of Payment Distribution]
+</p>
 
@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 4:Trends of Total Charges and Legnth of Stay </span>
+<p align = "center">
+[Figure 4:Trends of Total Charges and Legnth of Stay ]
+</p>
 
@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 5: Age Group Distribution of Patients in New York</span>
+<p align = "center">
+[Figure 5: Age Group Distribution of Patients in New York]
+</p>
 
@@
 
-## **Plot 6: Heatmap of Gender vs. Age Group**
-
-```{r}
-#| echo: false
-heatmap_data <- table(cleaned_data$`Age Group`, cleaned_data$Gender)
-
-row_names <- as.factor(rownames(heatmap_data))
-col_names <- as.factor(colnames(heatmap_data))
-
-heatmap(heatmap_data,
-    Rowv = NA,
-    Colv = NA,
-    col = colorRampPalette(c("white", "blue"))(50),
-    xlab = "Gender",
-    main = "Gender v. Age Group",
-    labRow = row_names,
-    labCol = col_names
-)
-
-```
-<span style="font-size: 14px; font-style: italic;">Figure 6: Gender vs Age Group Heatmapt</span>
-
-### Plot Desciption and Inferencing
-
-The following heatmap shows the relationship between the gender and age group for patients in New York in 2013.
-As we can observe, women tend to be patients at hospitals at a higher rate then men.
-Especially for the 70 or older bracket which is the highest frequency of hospital patients, females are significantly a larger proprotion of who are patients compared to men.
-Another important data point to understand is the identification of "U" or individuals who do not identify as male and female, which is high for ages 18-29, which is an important data point to help increase healthcare for individuals with fluid gender identity or may identify as non-binary.
-
-\newpage
-
-## Plot 7: Stacked Bar Chart by Gender and Race
+## Plot 6: Stacked Bar Chart by Gender and Race
 
@@
   scale_fill_manual(
-    values = c("White" = "lightblue", "African American" = "lightcoral", "Other" = "lightgreen"),
+    values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
     breaks = c("White", "Black", "Other"),
@@
 
-plot7 + ggtitle("Data Source: mtcars dataset, ggplot2 package")
+plot7 + ggtitle("Stacked Bar Chart by Gender and Race")
 
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 7: Stacked Bar Chart by Race and Gender</span>
+<p align = "center">
+[Figure 6: Stacked Bar Chart by Race and Gender]
+</p>
 
@@
 
-## Plot 8: Boxplots on the Total Costs on the Risk of Mortality
+## Plot 7: Boxplots on the Total Costs on the Risk of Mortality
 
@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 8: Boxplots on Total Cost and Mortality</span>
+<p align = "center">
+[Figure 7: Boxplots on Total Cost and Mortality]
+</p>
 
 ## Plot Description and Inferencing
-In Figure 8, our objective was to assess the central tendency of total hospital costs across different levels of Risk of Mortality.
+
+In Figure 7, our objective was to assess the central tendency of total hospital costs across different levels of Risk of Mortality.
 Employing separate box plots for each severity category allowed us to discern distinct patterns in the distribution of total costs associated with varying degrees of mortality risk.
-Notably, when focusing on Extreme cases of mortality, a clear trend emerged—median total costs were significantly higher, accompanied by a broader spread in the distribution.
+Notably, when focusing on Extreme cases of mortality, a clear trend emerged---median total costs were significantly higher, accompanied by a broader spread in the distribution.
 This observation implies that patients facing an Extreme risk of mortality not only incur higher median costs but also exhibit greater variability in the financial implications of their healthcare, suggesting a more diverse range of medical interventions and resource utilization.
@@
 Beyond its immediate implications, this nuanced analysis offers valuable insights for healthcare planning and resource allocation, emphasizing the need for tailored strategies based on the specific needs associated with varying risk levels.
-This approach contributes to more effective and efficient healthcare delivery, promoting a patient-centered and economically sustainable healthcare system.
 
@@
 
-## Plot 9: Violin Plot Comparing New York City and Western New York's Severity of Illness
+## Plot 8: Violin Plot Comparing New York City and Western New York's Severity of Illness
 
@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 9: NYC and Western NY Severity of Illness Comparsion </span>
+<p align = "center">
+[Figure 8: NYC and Western NY Severity of Illness Comparsion ]
+</p>
 
 ## Plot Description and Inferencing
-Figure 9 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
+
+Figure 8 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
 The focus is on the distribution of patients' APR severity codes, measured on a 1 to 4 scale (1 being mild, 2 being moderate, 3 being serious, and 4 being extreme).
@@
 
-## Plot 10: Distribution of APR Severity of Illness for Medicaid Patients
+## Plot 9: Distribution of APR Severity of Illness for Medicaid Patients
 
@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 10: APR Severity of Illness for Medicaid Patients </span>
+<p align = "center">
+[Figure 9: APR Severity of Illness for Medicaid Patients]
+</p>
 
 ## Plot Description and Inferencing
-In our examination of Figure 10, our focus shifted toward a specific subset of patients—those under Medicaid care.
+
+In our examination of Figure 9, our focus shifted toward a specific subset of patients---those under Medicaid care.
 Medicaid, a health program designed to support individuals classified as economically disadvantaged, prompted our investigation into the severity of diagnoses within this demographic.
@@
 
-## Plot 11: Facet Grid of Dotplots for Total Charges on Length of Stay
-
-```{r warning = FALSE}
-#| echo: false
-library(ggplot2)
-
-ggplot(cleaned_data, aes(x = cleaned_data$`Length of Stay`, y = cleaned_data$`Total Charges`)) +
-  geom_point(size = 0.2) +
-  facet_grid(Race ~ `Type of Admission`) +
-  labs(title = "Length of Stay v. Total Charges", x = "Length of Stay", y = "Total Charges") +
-  scale_y_continuous(labels = scales::comma, 
-                     breaks = scales::pretty_breaks(), 
-                     limits = c(0, max(cleaned_data$`Total Charges`)), # 
-                     expand = c(0, 0)) +
-  theme(axis.text = element_text(size = 8))
-
-```
-<span style="font-size: 14px; font-style: italic;">Figure 11: Dotplots for Total Charges on Length of Stay </span>
-
-## Plot Description and Inferencing
-
-In Figure 11, our aim was to dissect assessments based on both race and the type of admission.
-Through our analysis, a discernible pattern emerged, revealing consistent spikes in each category of total charges, particularly towards the lower end of the length of stay spectrum.
-This observation suggests that, irrespective of race, there are commonalities in the distribution of total charges, with pronounced concentration points occurring for shorter stays.
-
-Upon closer examination, the data points appear to be uniformly distributed across all racial categories, implying a degree of similarity in the cost distribution patterns among different racial groups.
-Interestingly, the primary differentiating factor contributing to variations in total costs is identified as the type of admission.
-This finding underscores the significance of admission characteristics in influencing healthcare costs, emphasizing that differences in total charges are more prominently linked to variations in admission types rather than racial distinctions.
-
-These insights shed light on the nuanced interplay between race, type of admission, and healthcare costs.
-Understanding these relationships is crucial for healthcare administrators and policymakers in designing targeted strategies for cost-effective care delivery, ensuring equitable access, and fostering a healthcare system that is both inclusive and economically sustainable.
-
-\newpage
-
-## Plot 12: Facet Density Plots of Total Cost per Age-Group Buckets
+## Plot 10: Facet Density Plots of Total Cost per Age-Group Buckets
 
@@
 ```
-<span style="font-size: 14px; font-style: italic;">Figure 12: Density Plots of Total Costs per Age-Groups</span>
+<p align = "center">
+[Figure 10: Density Plots of Total Costs per Age-Groups]
+</p>
 
 ## Plot Description and Inferencing
-In Figure 12, we analyzed age-specific total costs by charting density plots using a facet grid.
+
+In Figure 10, we analyzed age-specific total costs by charting density plots using a facet grid.
 For patients under 17, costs skewed right, suggesting lower expenses, potentially due to the overall good health of children.
@@
 ### References
+
 New York State Department of Health.

### 2023-10-31 DraftFour.qmd
Draft Four for real

--- 2023-10-08 NewSTAT405.qmd
@@
 
-*(last updated 10/07/2023)*
+*(last updated 10/30/2023)*
 
@@
 
-## **Plot 1: Price vs Length of Stay and Linear Regression Statistical Test**
-
-<br />
-
-```{r, include = FALSE}
+```{r, warning = FALSE, message = FALSE}
 #| echo: false
-# install.packages("readr")
-# install.packages("xtable")
+library(RSQLite)
 library(readr)
@@
 library(xtable)
-options(xtable.floating = FALSE)
-options(xtable.timestamp = "")
-data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv", show_col_types = FALSE)
 
-cleaned_data <- na.omit(data_set)
+dcon <- dbConnect(SQLite(), dbname = "finalDB.db", timeout = 10)
 
@@
 
-<br />
+## Plot 1: Price vs Length of Stay and Linear Regression Statistical Test
 
-```{r}
+```{r, results='asis', warning=FALSE}
 #| echo: false
-y <- cleaned_data$`Total Charges`
-```
-
-```{r, results='asis'}
-#| echo: false
-length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
+query <- dbSendQuery(dcon,"
+SELECT *
+FROM '2013data';
+")
+cleaned_data <- dbFetch(query, -1)
+dbClearResult(query)
+y <- cleaned_data$`TotalCharges`
+length_of_stay <- strtoi(cleaned_data$`LengthofStay`)
 
@@
 ```
-<p align = "center">
-[Figure 1a: Length of Stay on Price Regression Dotplot]
-</p>
 
+```{=tex}
+\begin{center}
+Figure 1a: Length of Stay on Price Regression Dotplot
+\end{center}
+```
 ```{r, results='asis'}
@@
 ```
-<p align = "center">
-[Figure 1b: Table of Linear Regression Modeling Data]
-</p>
+
+Table 1: Table of Linear Regression Modeling Data
 
@@
 
-### Plot Desciption and Inferencing
+\newpage
 
-In addition to the regression analysis, we visualized the relationship between the length of stay and the associated price using a scatter plot with a fitted regression line.
-The scatter plot provides a clear graphical representation of the positive correlation between these variables.
-As the length of stay increases, there is a discernible upward trend in price.
-The regression line, which is based on the linear regression model we performed, further accentuates this trend.
-It serves as a useful tool for illustrating the direction and strength of the relationship, making it easier to interpret the practical implications of our findings.
-This visualization reinforces the statistical significance we observed in our analysis and enhances our understanding of the relationship between patient length of stay and healthcare costs for patients in the state of New York in 2013.
+## Plot 2: Heat Map of Individuals on Medicaid by Race and Borough**
+
+```{r, fig.height = 6, fig.width = 9, message = FALSE, warning = FALSE}
+#| echo: false
+library(DBI)
+library(RSQLite)
+library(ggplot2)
+library(sf)
+library(tigris)
+#| echo: false
+
+query <- dbSendQuery(dcon, "
+  SELECT `HospitalCounty`, Race, COUNT(*) as count
+  FROM `2013data`
+  WHERE `HealthServiceArea` = 'New York City' 
+    AND Race IN ('Black/African American', 'White')
+  GROUP BY `HospitalCounty`, `Race`
+")
+
+data <- dbFetch(query, -1)
+
+ggplot(data, aes(x = `HospitalCounty`, y = `Race`, fill = count)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "red") +
+  labs(title = "Number of Black/Other Race Individuals on Medicaid in NYC", 
+       x = "Borough", 
+       y = "Race", 
+       fill = "Count") +
+  theme_minimal()
+
+dbClearResult(query)
+
+# Query the data
+query <- dbSendQuery(dcon,"
+  SELECT `HospitalCounty`, Race, COUNT(*) as count
+  FROM `2013data`
+  WHERE `HealthServiceArea` = 'New York City' 
+    AND Race IN ('Black/African American', 'White')
+  GROUP BY `HospitalCounty`, `Race`
+"
+)
+data <- dbFetch(query, -1)
+
+dbClearResult(query)
+
+# Retrieve the NYC boroughs shape data and filter for the five boroughs
+options(tigris_use_cache = TRUE)
+
+# Retrieve the NYC boroughs shape data and filter for the five boroughs
+ny_counties <- tigris::counties(state = "NY", cb = TRUE) %>%
+  st_as_sf()
+
+nyc_boroughs <- ny_counties[ny_counties$NAME %in% c("Bronx", "Kings", "Manhattan", "Queens", "Richmond"), ]
+
+# Merge the data with the nyc_boroughs spatial object
+data_sf <- merge(nyc_boroughs, data, by.x = "NAME", by.y = "HospitalCounty")
+
+# Compute the centroids for the NYC boroughs for plotting circles
+data_sf$centroid <- st_centroid(st_geometry(data_sf))
+```
+
+```{=tex}
+\begin{center}
+Figure 2: Individuals on Medicaid by Race and Borough
+\end{center}
+```
+\newpage
+## Plot 3: Map of Individuals on Medicare by Race**
+
+```{r, warning = FALSE}
+#| echo: false
+# Plotting side-by-side maps
+ggplot(data_sf) +
+  geom_sf(aes(fill = NAME), color = "grey60", size = 0.5) +  # Borough fill
+  geom_point(aes(x = st_coordinates(centroid)[, "X"], 
+                 y = st_coordinates(centroid)[, "Y"], 
+                 size = count, 
+                 color = Race), 
+             alpha = 0.7, show.legend = "point") +  # Plot circles based on count, color-coded by Race
+
+  # Faceting for each race
+  facet_wrap(~Race, ncol = 2) +
+
+  # Custom color scale
+  scale_color_manual(values = c("Black/African American" = "blue", "White" = "red"), name = "Race") +
+  scale_size_continuous(name = "Count", breaks = c(500, 1000, 1500, 2000)) +  # Adjust the size scale for circles
+
+  # Title and subtitles
+  labs(title = "Number of Black vs. White Individuals on Medicaid in NYC Boroughs",
+       subtitle = "Size of point indicates count, color indicates race") +
+
+  # Minimal theme and customization
+  theme_minimal() +
+  theme(legend.position = "right",
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank())
+
+```
+
+```{=tex}
+\begin{center}
+Figure 3: Map of NYC displaying number of individuals on Medicaid by Borough
+\end{center}
+```
+## Plot Description and Inferencing
 
@@
 
-## **Plot 2: Boxplot of Total Costs**
+## Plot 4: Waffle Plot of Severity of Illness by Race**
 
@@
 #| echo: false
-cleaned_data$"Total Costs"[cleaned_data$"Total Costs" > quantile(cleaned_data$"Total Costs", 0.95)] <- quantile(cleaned_data$"Total Costs", 0.95)
+library(waffle)
+library(patchwork)
+query <- dbSendQuery(dcon,"
+SELECT Race, APRSeverityofIllnessCode FROM '2013data';
+                     ")
+IllnessCodes <- dbFetch(query, -1)
+dbClearResult(query)
 
-#summary(cleaned_data$"Total Costs")
+waffle_data1 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "Black/African American"])/2500)
+waffle_data2 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "White"])/5000)
 
-boxplot(cleaned_data$"Total Costs",
-    main = "Box Plot of Total Costs (Outliers Truncated)",
-    ylab = "Total Costs",
-    horizontal = TRUE
-)
+layout(matrix(c(1, 2), nrow = 1))
+
+plot1 <- waffle(waffle_data1, rows = 8, size = .5, 
+            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "Black/African Americans")
+plot2 <- waffle(waffle_data2, rows = 8, size = .5, 
+            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "White")
+
+arrange.plots <- plot1 + theme(plot.title = element_text(size = 10)) + plot2 + theme(plot.title = element_text(size = 10))
+arrange.plots
+```
+\begin{center}
+Figure 4: Waffle Plots Depicting Frequencies of APR Codes 
+\end{center}
+
+## Plot Description and Inferencing
+In the preceding graphical representation, we have visualized the distribution of patients based on their racial backgrounds across the various APR Severity of Illness Codes.
+These codes are categorized on a scale from 1 to 4, with higher values indicating increased severity.
+The observed trend indicates that patients of White ethnicity tend to be associated with higher severity APR codes, potentially implying a discrepancy in the prioritization of patient care between White and African American populations.
+
+\newpage
+## Plot 5: Barplots of Average Total Charges by Race 2013 vs 2021
+
+```{r, fig.height = 7, fig.width = 5}
+#| echo: false
+library(ggplot2)
+library(patchwork)
+
+pregnant_womenQ <- dbSendQuery(dcon, "
+    SELECT Race, TotalCharges
+    FROM '2013data'
+    WHERE APRMDCCode == 14 OR APRMDCCode == 15;")
+pregnant_womenDF = dbFetch(pregnant_womenQ, -1)
+dbClearResult(pregnant_womenQ)
+
+plot_race2013 <- ggplot(pregnant_womenDF) + 
+  aes(x = Race, y = TotalCharges) +
+  stat_summary(fun = "mean", geom = "bar", fill = "salmon") +  # Calculate and plot averages
+  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2013") +  # Customize axis labels
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size = 7))  # Rotate x-axis labels for
+
+pregnant_womenQ2 <- dbSendQuery(dcon, "
+    SELECT Race, TotalCharges
+    FROM '2021data'
+    WHERE APRMDCCode == 14 OR APRMDCCode == 15 AND Race != 'Multi-racial';")
+pregnant_womenDF2 = dbFetch(pregnant_womenQ2, -1)
+dbClearResult(pregnant_womenQ2)
+
+plot_race2021 <- ggplot(pregnant_womenDF2) + 
+  aes(x = Race, y = TotalCharges) +
+  stat_summary(fun = "mean", geom = "bar", fill = "violet") +  # Calculate and plot averages
+  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2021") +  # Customize axis labels
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size = 7))  # Rotate x-axis labels for
+
+combined_plot_race <- plot_race2013 + plot_race2021
+combined_plot_race
 
 ```
-<p align = "center">
-[Figure 2: Total Costs Boxplot]
-</p>
-
-### Plot Desciption and Inferencing
-
-The following boxplot refers to our Total Cost data and is representative of our data that has been cleaned to remove outliers that are in our top 5% of data points, meaning that the outliers that are extremely high are cleaned out.
-In our analysis, we used a horizontal box plot to summarize the distribution of 'Total Costs' in our dataset.
-The median, located at approximately \$8,247.96, divides the data into two halves, with the interquartile range (IQR) spanning from Q1 at \$4,519.54 to Q3 at \$15,840.81.
-Outliers, representing data points significantly deviating from the central distribution, are visible.
-The total cost range extends from \$0.41 to \$45,292.36, showcasing the data's spread.
-The mean total cost is around \$12,360.20, but outliers can influence this value.
-The boxplot also is skewed to the right.
-This visualization provides key insights into our cost structure analysis.
-
-From this we can infer that there are many outliers, however, in the discussion of costs they are heavily spread and that most patients are paying within our IQR but there are many different outliers that spread when we are discussing costs for healthcare.
-
+\begin{center}
+Figure 5: Barplots of Average Total Charges by Race in 2013 and 2021
+\end{center}
+## Plot Description and Inferencing
+In figure 5, the bar plot illustrates a compelling contrast in average total charges for hospital care among different racial groups.
+Notably, it highlights that individuals who identify as Black and those categorized under the "Other Races" group experience significantly higher average total charges compared to their White counterparts.
+This disparity in healthcare costs sheds light on the existing inequalities within the healthcare system, prompting a critical examination of the factors contributing to these disparities and the potential need for targeted interventions to address them.
+We see that in 2021, post the COVID-19
+pandemic, this disparity between average total charges of Black patients and White patients
+increased, showing that the pandemic potentially increased already present disparity in healthcare.
 \newpage
-
-## **Plot 3: Payment Type Distribution**
-
-```{r, warning = FALSE, fig.height = 15, fig.width = 50}
-#| echo: false
-# Set the margin to include more space at the bottom
-par(mar = c(5, 4, 4, 2) + 0.1)
-par(cex.main = 5 )
-cleaned_data$`Payment Typology 1` <- as.factor(cleaned_data$`Payment Typology 1`)
-payments <- table(cleaned_data$`Payment Typology 1`)
-pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA)
-legend(1, 1, 1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 3)
-
-```
-<p align = "center">
-[Figure 3: Pie Chart of Payment Distribution]
-</p>
-
-### Plot Desciption and Inferencing
-
-The following piechart identifies the payment type distribution for patients in New York in 2013.
-As we can observe, most patients are using Medicare as a way to pay for their medical bills followed by Medicaid.
-These government programs are important to New York residents and provides important data on the impact of healthcare programs.
-The exact percentagse are Medicaid at 59.7% and Medicaid at 26.7%, followed by Private Insurance at around 10%.
-This information reveals how important the impact of governmental insurance programs are to individuals in New York.
-
-\newpage
-
-## **Plot 4: Trends in Total Charges and Length of Stay**
-
-```{r warning = FALSE}
-#| echo: false
-
-plot(cleaned_data$`Length of Stay`, cleaned_data$`Total Charges`, type = "l", col = "blue", xlab = "Length of Stay", ylab = "Total Charges", main = "Trends in Total Charges and Length of Stay")
-
-lines(cleaned_data$`Birth Weight`, cleaned_data$`Total Costs`, col = "red")
-
-legend("topright", legend = c("Total Charges", "Total Costs"), col = c("blue", "red"), lty = 1)
-
-grid()
-```
-<p align = "center">
-[Figure 4:Trends of Total Charges and Legnth of Stay ]
-</p>
-
-### Plot Desciption and Inferencing
-
-The plotted data and corresponding projections unmistakably depict a consistent total cost profile, irrespective of variations in both the length of stay and total charges.
-This apparent lack of correlation between the length of stay and associated costs prompts a deeper inquiry into the potential contributing factors.
-It would be important to explore the influence of fixed operational costs, negotiated contractual agreements, and astute cost management strategies that may explain this observed financial consistancy between the costs for a hospital and for its patients
-
-\newpage
-
-## **Plot 5: Barplot of Age Group Distribution of Patients**
-
-```{r}
-#| echo: false
-ageGroups <- table(cleaned_data$`Age Group`)
-barplot(ageGroups,
-    main = "Age Group Distribution", xlab = "Age Group",
-    ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)
-)
-
-```
-<p align = "center">
-[Figure 5: Age Group Distribution of Patients in New York]
-</p>
-
-### Plot Desciption and Inferencing
-
-Another important piece of data that we wanted to observe is the amount of patients from each age group of the population.
-The data is divided into five segmests 0-17, 18-29, 30-49, 50-69, and 70-older.
-As we can see there is an upward trend of the amount of patients as the ages of the patients increases, although there is a small decrease between 0-17 and 8-29.
-The largest frequency if the 70-older age bucket which has around 289248 compared to 18-29 with our smallest frequency of 53009.This increase in older patients reveals the impact of older populations in New York and the necesity to increase health care efforts towards older populations.
-
-\newpage
-
-## Plot 6: Stacked Bar Chart by Gender and Race
-
-```{r}
-#| echo: false
-# install.packages("ggplot2")
-library(ggplot2)
-
-plot7 <- ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
-  geom_bar() +
-  labs(
-    title = "Stacked Bar Chart by Gender and Race",
-    x = "Gender",
-    y = "Count"
-  ) +
-  scale_fill_manual(
-    values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
-    breaks = c("White", "Black", "Other"),
-    labels = c("White", "Black", "Other"),
-    name = "Race"
-  ) +
-  theme_minimal()
-
-plot7 + ggtitle("Stacked Bar Chart by Gender and Race")
-
-```
-<p align = "center">
-[Figure 6: Stacked Bar Chart by Race and Gender]
-</p>
-
-## Plot Description and Inferencing
-
-In the analysis of patient demographics in the year 2013, a bar chart was employed to visually represent the breakdown of patients by gender and further stratify each gender by race.
-A notable observation from the chart is the significant predominance of female patients over male patients during this period.
-
-Upon closer examination of the male demographic, the distribution reveals a relatively balanced representation between individuals identified as white and those classified as not white.
-This suggests a relatively even distribution among male patients concerning racial categories.
-In contrast, the female demographic exhibits a slightly more diverse distribution across racial categories.
-This observation implies that, among female patients, there is a greater degree of heterogeneity in terms of racial composition compared to their male counterparts.
-
-These findings contribute valuable insights into the gender and racial composition of the patient population in 2013, emphasizing the need for a nuanced understanding of demographic trends within healthcare settings.
-Further exploration and analysis of these demographic patterns may reveal additional insights relevant to healthcare planning, resource allocation, and the provision of patient-centered care.
-
-\newpage
-
-## Plot 7: Boxplots on the Total Costs on the Risk of Mortality
+## Plot 6: Boxplots on the Total Costs on the Risk of Mortality
 
@@
 ```
-<p align = "center">
-[Figure 7: Boxplots on Total Cost and Mortality]
-</p>
 
+```{=tex}
+\begin{center}
+Figure 6: Boxplots on Total Cost and Mortality
+\end{center}
+```
 ## Plot Description and Inferencing
 
-In Figure 7, our objective was to assess the central tendency of total hospital costs across different levels of Risk of Mortality.
+In Figure 6, our objective was to assess the central tendency of total hospital costs across different levels of Risk of Mortality.
 Employing separate box plots for each severity category allowed us to discern distinct patterns in the distribution of total costs associated with varying degrees of mortality risk.
@@
 
-## Plot 8: Violin Plot Comparing New York City and Western New York's Severity of Illness
+## Plot 7: Violin Plot Comparing New York City and Western New York's Severity of Illness
 
@@
 ```
-<p align = "center">
-[Figure 8: NYC and Western NY Severity of Illness Comparsion ]
-</p>
+
+\begin{center}
+Figure 7: NYC and Western NY Severity of Illness Comparsion
+\end{center}
 
@@
 
-Figure 8 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
+Figure 7 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
 The focus is on the distribution of patients' APR severity codes, measured on a 1 to 4 scale (1 being mild, 2 being moderate, 3 being serious, and 4 being extreme).
@@
 
-## Plot 9: Distribution of APR Severity of Illness for Medicaid Patients
+## Plot 8: Distribution of APR Severity of Illness for Medicaid Patients
 
@@
 ```
-<p align = "center">
-[Figure 9: APR Severity of Illness for Medicaid Patients]
-</p>
 
+\begin{center}
+Figure 8: APR Severity of Illness for Medicaid Patients
+\end{center}
 ## Plot Description and Inferencing
 
-In our examination of Figure 9, our focus shifted toward a specific subset of patients---those under Medicaid care.
+In our examination of Figure 8, our focus shifted toward a specific subset of patients---those under Medicaid care.
 Medicaid, a health program designed to support individuals classified as economically disadvantaged, prompted our investigation into the severity of diagnoses within this demographic.
@@
 
-## Plot 10: Facet Density Plots of Total Cost per Age-Group Buckets
+## Plot 9: Facet Density Plots of Total Cost per Age-Group Buckets
 
@@
 ```
-<p align = "center">
-[Figure 10: Density Plots of Total Costs per Age-Groups]
-</p>
 
+```{=tex}
+\begin{center}
+
+Figure 9: Density Plots of Total Costs per Age-Groups
+
+\end{center}
+```
 ## Plot Description and Inferencing
 
-In Figure 10, we analyzed age-specific total costs by charting density plots using a facet grid.
+In Figure 9, we analyzed age-specific total costs by charting density plots using a facet grid.
 For patients under 17, costs skewed right, suggesting lower expenses, potentially due to the overall good health of children.
@@
 
+## Plot 10: Pie Charts of Payment Type by Race
+```{r}
+#| echo: false
+library(ggplot2)
+library(gridExtra)
+library(patchwork)
+
+paymentTypeAndRaceQ <- dbSendQuery(dcon, "
+                                    SELECT Race, PaymentTypology1 FROM '2013data' WHERE Race != 'Unknown';")
+paymentTypeAndRaceDF <- dbFetch(paymentTypeAndRaceQ, -1)
+dbClearResult(paymentTypeAndRaceQ)
+str(paymentTypeAndRaceDF)
+
+pie_charts <- lapply(unique(paymentTypeAndRaceDF$Race), function(race) {
+  data_subset <- subset(paymentTypeAndRaceDF, Race == race)
+  pie_chart <- ggplot(data_subset, aes(x = "", fill = PaymentTypology1)) +
+    geom_bar(width = 1) +
+    coord_polar("y") +
+    ggtitle(race) +
+    theme_void() +
+    scale_fill_brewer(palette = "Set3") +
+    theme(legend.position = ifelse(race == 'Black/African American', "left", "none"),
+          axis.text = element_text(size = 0),
+          legend.text = element_text(size = 5)) +
+    labs(fill = "Payment Type")
+
+  return(pie_chart)
+})
+pie_charts[[1]] + pie_charts[[2]] + pie_charts[[3]] + plot_layout(ncol = 1)
+```
+\begin{center}
+Figure 10: Pie Charts on Payment Type by Race
+\end{center}
+
+## Plot Description and Inferencing
+
 ### References

### 2023-10-31 DraftFour.qmd
Update DraftFour.qmd

@@
 
+Pie Chart 1: White Patients
+
+In the first pie chart, which represents White patients, a significant portion of them use Medicare as their Payment Type.
+This choice reflects a trend among White patients, which is that a high percentage of them will rely on Medicare to cover their Hospital Charges.
+Between those that are not using Medicare, they would instead use Medicaid, Blue Cross/Blue Shield, or Private Health Insurances.
+
+Pie Chart 2: Black/African American Patients
+
+The second pie chart, which illustrates Black/African American patients, reveals a inclination towards Medicare, but a significant amount of patients would also use Medicaid.
+This suggests that there is a dual reliance on both the Medicare and Medicaid insurances, and this pie chart shows this trend.
+
+Pie Chart 3: Other Race Patients
+
+In the third pie chart, which represents the patients that are of other races, we see a greater split.
+Approximately half of the patients choose Medicare, and a majority of the patients that have not chosen Medicare rely on Medicaid instead.
+This shows a somewhat balanced usage between Medicare and Medicaid by this demographic of patients.
+
+These insights, derived from the pie charts, shed light on the nuanced payment patterns across different racial groups in the context of the year 2013, emphasizing the significance of understanding and catering to diverse payment preferences within the healthcare system of that time.
 ### References

### 2023-10-31 DraftFour.qmd
Update DraftFour.qmd

--- 2023-10-31 DraftFour.qmd
@@
 library(xtable)
+library(ggplot2)
 
@@
 #| echo: false
-# install.packages("ggplot2")
+library(ggplot2)
 ggplot(cleaned_data) +
-  aes(x = `APR Risk of Mortality`, y = `Total Costs`) +
+  aes(x = `APRRiskofMortality`, y = `TotalCosts`) +
   geom_boxplot(color = "salmon") +
@@
 #| echo: false
-subsetted_data <- cleaned_data[cleaned_data$`Health Service Area` %in% c("New York City", "Western NY"), ]
+subsetted_data <- cleaned_data[cleaned_data$`HealthServiceArea` %in% c("New York City", "Western NY"), ]
 
@@
 # Create a violin plot for APR Severity of Illness Code
-ggplot(subsetted_data, aes(x = factor(`Health Service Area`), y = `APR Severity of Illness Code`, fill = `Health Service Area`)) +
+ggplot(subsetted_data, aes(x = factor(`HealthServiceArea`), y = `APRSeverityofIllnessCode`, fill = `HealthServiceArea`)) +
   geom_violin() +
@@
 medicaid_data <- cleaned_data %>%
-  filter(`Payment Typology 1` == "Medicaid")
+  filter(`PaymentTypology1` == "Medicaid")
 
 # Create a ridgeline plot to visualize APR Severity of Illness
-ggplot(medicaid_data, aes(x = `APR Severity of Illness Code`, y = ..density.., fill = `APR Severity of Illness Code`)) +
+ggplot(medicaid_data, aes(x = `APRSeverityofIllnessCode`, y = ..density.., fill = `APRSeverityofIllnessCode`)) +
   geom_density(alpha = 0.6) +
@@
 #| echo: false
-gg <- ggplot(cleaned_data, aes(x = `Total Costs`, y = ..density.., fill = `Age Group`)) +
+gg <- ggplot(cleaned_data, aes(x = `TotalCosts`, y = ..density.., fill = `AgeGroup`)) +
   geom_density(alpha = 0.5) +
-  facet_grid(`Age Group` ~ .) +
+  facet_grid(`AgeGroup` ~ .) +
   labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density")

### 2023-10-31 DraftFour.qmd
Update DraftFour.qmd

@@
 ```
-
+\begin{center}
 Table 1: Table of Linear Regression Modeling Data
+\end{center}
 
@@
 ```
+## Plot Description and Inferencing
+
 \newpage
@@
 ## Plot Description and Inferencing
+Pie Chart 1: White Patients
+
+In the first pie chart, which represents White patients, a significant portion of them use Medicare as their Payment Type.
+This choice reflects a trend among White patients, which is that a high percentage of them will rely on Medicare to cover their Hospital Charges.
+Between those that are not using Medicare, they would instead use Medicaid, Blue Cross/Blue Shield, or Private Health Insurances.
+
+Pie Chart 2: Black/African American Patients
+
+The second pie chart, which illustrates Black/African American patients, reveals a inclination towards Medicare, but a significant amount of patients would also use Medicaid.
+This suggests that there is a dual reliance on both the Medicare and Medicaid insurances, and this pie chart shows this trend.
+
+Pie Chart 3: Other Race Patients
+
+In the third pie chart, which represents the patients that are of other races, we see a greater split.
+Approximately half of the patients choose Medicare, and a majority of the patients that have not chosen Medicare rely on Medicaid instead.
+This shows a somewhat balanced usage between Medicare and Medicaid by this demographic of patients.
+
+These insights, derived from the pie charts, shed light on the nuanced payment patterns across different racial groups in the context of the year 2013, emphasizing the significance of understanding and catering to diverse payment preferences within the healthcare system of that time.
 

### 2023-10-31 DraftFour.qmd
Update DraftFour.qmd

@@
 library(xtable)
-library(ggplot2)
 
@@
 \begin{center}
-Figure 1a: Length of Stay on Price Regression Dotplot
+Figure 1: Length of Stay on Price Regression Dotplot
 \end{center}
@@
 ```
-\begin{center}
+
 Table 1: Table of Linear Regression Modeling Data
-\end{center}
 
@@
 
-## Plot 2: Heat Map of Individuals on Medicaid by Race and Borough**
+## Plot 2: Heat Map of Individuals on Medicaid by Race and Borough
 
@@
 ## Plot Description and Inferencing
+Figure 2 is a heatmap illustrating the distribution of individuals on Medicaid across New York City boroughs, segmented by race.
+The X-axis represents boroughs such as Bronx, Kings (Brooklyn), Manhattan, Queens, and Richmond (Staten Island), while the Y-axis distinguishes between White and Black/African American racial groups.
+The color gradient, ranging from light to dark red, indicates the count of individuals, with darker tones signifying higher numbers.
+Notably, Kings (Brooklyn) has a significant number of Black/African American individuals on Medicaid, while Manhattan displays a higher count of White individuals.
+Richmond (Staten Island) exhibits lower numbers for both racial categories.
+Meanwhile, Bronx and Queens show a relatively balanced distribution between the two groups.
+This visualization offers essential insights into the racial demographics of Medicaid beneficiaries in NYC, serving as a valuable resource for policymakers and healthcare administrators.
 
 \newpage
-## Plot 3: Map of Individuals on Medicare by Race**
+## Plot 3: Map of Individuals on Medicare by Race
 
@@
 ## Plot Description and Inferencing
+Figure 3 presents a geographical representation of New York City, displaying the number of individuals on Medicaid across different boroughs, differentiated by race.
+The boroughs—Bronx, Kings (Brooklyn), Queens, and Richmond (Staten Island)—are distinctly colored, while two circular points within each borough indicate the counts of Black/African American and White individuals.
+The size of the circles signifies the number of individuals, with larger circles representing higher counts, and the color of each circle (blue for Black/African American and red for White) represents the respective racial groups.
+A notable observation is the prominent size of the blue circle in Kings (Brooklyn), highlighting a substantial count of Black/African American individuals on Medicaid.
+Conversely, Manhattan has a noticeably larger red circle, indicating a higher number of White individuals.
+This map offers a visual synthesis of Medicaid beneficiaries' racial distribution across NYC boroughs, informing stakeholders about the demographic landscape of healthcare recipients in the metropolis.
 
@@
 
-## Plot 4: Waffle Plot of Severity of Illness by Race**
+## Plot 4: Waffle Plot of Severity of Illness by Race
 
@@
 \newpage
-## Plot 6: Boxplots on the Total Costs on the Risk of Mortality
 
-```{r warning = FALSE}
-#| echo: false
-library(ggplot2)
-ggplot(cleaned_data) +
-  aes(x = `APRRiskofMortality`, y = `TotalCosts`) +
-  geom_boxplot(color = "salmon") +
-  labs(
-    title = "Total Costs on Risk of Mortality",
-    x = "APR Risk of Mortality",
-    y = "Total Costs"
-  ) +
-  theme_minimal()
-
-```
-
-```{=tex}
-\begin{center}
-Figure 6: Boxplots on Total Cost and Mortality
-\end{center}
-```
-## Plot Description and Inferencing
-
-In Figure 6, our objective was to assess the central tendency of total hospital costs across different levels of Risk of Mortality.
-Employing separate box plots for each severity category allowed us to discern distinct patterns in the distribution of total costs associated with varying degrees of mortality risk.
-Notably, when focusing on Extreme cases of mortality, a clear trend emerged---median total costs were significantly higher, accompanied by a broader spread in the distribution.
-This observation implies that patients facing an Extreme risk of mortality not only incur higher median costs but also exhibit greater variability in the financial implications of their healthcare, suggesting a more diverse range of medical interventions and resource utilization.
-
-This finding aligns with our initial expectation that higher levels of risk of mortality correspond to elevated costs of care.
-The discernible trend in the box plots reinforces the intuitive understanding that patients with a greater severity of illness often require more extensive and resource-intensive medical interventions, leading to increased healthcare costs.
-Beyond its immediate implications, this nuanced analysis offers valuable insights for healthcare planning and resource allocation, emphasizing the need for tailored strategies based on the specific needs associated with varying risk levels.
-
-\newpage
-
-## Plot 7: Violin Plot Comparing New York City and Western New York's Severity of Illness
+## Plot 6: Violin Plot Comparing New York City and Western New York's Severity of Illness
 
@@
 \begin{center}
-Figure 7: NYC and Western NY Severity of Illness Comparsion
+Figure 6: NYC and Western NY Severity of Illness Comparsion
 \end{center}
@@
 
-Figure 7 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
+Figure 6 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
 The focus is on the distribution of patients' APR severity codes, measured on a 1 to 4 scale (1 being mild, 2 being moderate, 3 being serious, and 4 being extreme).
@@
 
-## Plot 8: Distribution of APR Severity of Illness for Medicaid Patients
+## Plot 7: Distribution of APR Severity of Illness for Medicaid Patients
 
@@
 \begin{center}
-Figure 8: APR Severity of Illness for Medicaid Patients
+Figure 7: APR Severity of Illness for Medicaid Patients
 \end{center}
@@
 
-In our examination of Figure 8, our focus shifted toward a specific subset of patients---those under Medicaid care.
+In our examination of Figure 7, our focus shifted toward a specific subset of patients---those under Medicaid care.
 Medicaid, a health program designed to support individuals classified as economically disadvantaged, prompted our investigation into the severity of diagnoses within this demographic.
@@
 
-## Plot 9: Facet Density Plots of Total Cost per Age-Group Buckets
+## Plot 8: Facet Density Plots of Total Cost per Age-Group Buckets
 
-```{r warning = FALSE}
+```{r warning = FALSE, fig.width = 7, fig.height = 5}
 #| echo: false
-gg <- ggplot(cleaned_data, aes(x = `TotalCosts`, y = ..density.., fill = `AgeGroup`)) +
+data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)
+
+clean <- na.omit(data_set)
+
+gg <- ggplot(clean, aes(x = `Total Costs`, y = ..density.., fill = `Age Group`)) +
   geom_density(alpha = 0.5) +
-  facet_grid(`AgeGroup` ~ .) +
+  facet_grid(`Age Group` ~ .) +
   labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density")
 
-gg + scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
+gg + 
+  scale_x_continuous(limits = c(0, 35000)) +
+  scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
   labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density (in thousands)") +
@@
 
-Figure 9: Density Plots of Total Costs per Age-Groups
+Figure 8: Density Plots of Total Costs per Age-Groups
 
@@
 
-In Figure 9, we analyzed age-specific total costs by charting density plots using a facet grid.
+In Figure 8, we analyzed age-specific total costs by charting density plots using a facet grid.
 For patients under 17, costs skewed right, suggesting lower expenses, potentially due to the overall good health of children.
@@
 
-## Plot 10: Pie Charts of Payment Type by Race
-```{r}
+## Plot 9: Pie Charts of Payment Type by Race
+```{r, message = FALSE}
 #| echo: false
@@
 dbClearResult(paymentTypeAndRaceQ)
-str(paymentTypeAndRaceDF)
+#str(paymentTypeAndRaceDF)
 
@@
 ```
+
+```{=tex}
 \begin{center}
-Figure 10: Pie Charts on Payment Type by Race
+Figure 9: Pie Charts on Payment Type by Race
 \end{center}
+```
 
 ## Plot Description and Inferencing
+
 Pie Chart 1: White Patients
@@
 
+\newpage
+
 ### References

### 2023-11-07 NewSTAT405.qmd
Update NewSTAT405.qmd
Bertolusso fr

@@
 ```
-<p align = "center">
-[Figure 1a: Length of Stay on Price Regression Dotplot]
-</p>
 
+```{=tex}
+\begin{center}
+Figure 1: Length of Stay on Price Regression Dotplot
+\end{center}
+```
 ```{r, results='asis'}
@@
 
-### Plot Desciption and Inferencing
+\newpage
 
-In addition to the regression analysis, we visualized the relationship between the length of stay and the associated price using a scatter plot with a fitted regression line.
-The scatter plot provides a clear graphical representation of the positive correlation between these variables.
-As the length of stay increases, there is a discernible upward trend in price.
-The regression line, which is based on the linear regression model we performed, further accentuates this trend.
-It serves as a useful tool for illustrating the direction and strength of the relationship, making it easier to interpret the practical implications of our findings.
-This visualization reinforces the statistical significance we observed in our analysis and enhances our understanding of the relationship between patient length of stay and healthcare costs for patients in the state of New York in 2013.
+## Plot 2: Heat Map of Individuals on Medicaid by Race and Borough
+
+```{r, fig.height = 6, fig.width = 9, message = FALSE, warning = FALSE}
+#| echo: false
+library(DBI)
+library(RSQLite)
+library(ggplot2)
+library(sf)
+library(tigris)
+#| echo: false
+
+query <- dbSendQuery(dcon, "
+  SELECT `HospitalCounty`, Race, COUNT(*) as count
+  FROM `2013data`
+  WHERE `HealthServiceArea` = 'New York City' 
+    AND Race IN ('Black/African American', 'White')
+  GROUP BY `HospitalCounty`, `Race`
+")
+
+data <- dbFetch(query, -1)
+
+ggplot(data, aes(x = `HospitalCounty`, y = `Race`, fill = count)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "red") +
+  labs(title = "Number of Black/Other Race Individuals on Medicaid in NYC", 
+       x = "Borough", 
+       y = "Race", 
+       fill = "Count") +
+  theme_minimal()
+
+dbClearResult(query)
+
+# Query the data
+query <- dbSendQuery(dcon,"
+  SELECT `HospitalCounty`, Race, COUNT(*) as count
+  FROM `2013data`
+  WHERE `HealthServiceArea` = 'New York City' 
+    AND Race IN ('Black/African American', 'White')
+  GROUP BY `HospitalCounty`, `Race`
+"
+)
+data <- dbFetch(query, -1)
+
+dbClearResult(query)
+
+# Retrieve the NYC boroughs shape data and filter for the five boroughs
+options(tigris_use_cache = TRUE)
+
+# Retrieve the NYC boroughs shape data and filter for the five boroughs
+ny_counties <- tigris::counties(state = "NY", cb = TRUE) %>%
+  st_as_sf()
+
+nyc_boroughs <- ny_counties[ny_counties$NAME %in% c("Bronx", "Kings", "Manhattan", "Queens", "Richmond"), ]
+
+# Merge the data with the nyc_boroughs spatial object
+data_sf <- merge(nyc_boroughs, data, by.x = "NAME", by.y = "HospitalCounty")
+
+# Compute the centroids for the NYC boroughs for plotting circles
+data_sf$centroid <- st_centroid(st_geometry(data_sf))
+```
+
+```{=tex}
+\begin{center}
+Figure 2: Individuals on Medicaid by Race and Borough
+\end{center}
+```
+## Plot Description and Inferencing
+Figure 2 is a heatmap illustrating the distribution of individuals on Medicaid across New York City boroughs, segmented by race.
+The X-axis represents boroughs such as Bronx, Kings (Brooklyn), Manhattan, Queens, and Richmond (Staten Island), while the Y-axis distinguishes between White and Black/African American racial groups.
+The color gradient, ranging from light to dark red, indicates the count of individuals, with darker tones signifying higher numbers.
+Notably, Kings (Brooklyn) has a significant number of Black/African American individuals on Medicaid, while Manhattan displays a higher count of White individuals.
+Richmond (Staten Island) exhibits lower numbers for both racial categories.
+Meanwhile, Bronx and Queens show a relatively balanced distribution between the two groups.
+This visualization offers essential insights into the racial demographics of Medicaid beneficiaries in NYC, serving as a valuable resource for policymakers and healthcare administrators.
+
+\newpage
+## Plot 3: Map of Individuals on Medicare by Race
+
+```{r, warning = FALSE}
+#| echo: false
+# Plotting side-by-side maps
+ggplot(data_sf) +
+  geom_sf(aes(fill = NAME), color = "grey60", size = 0.5) +  # Borough fill
+  geom_point(aes(x = st_coordinates(centroid)[, "X"], 
+                 y = st_coordinates(centroid)[, "Y"], 
+                 size = count, 
+                 color = Race), 
+             alpha = 0.7, show.legend = "point") +  # Plot circles based on count, color-coded by Race
+
+  # Faceting for each race
+  facet_wrap(~Race, ncol = 2) +
+
+  # Custom color scale
+  scale_color_manual(values = c("Black/African American" = "blue", "White" = "red"), name = "Race") +
+  scale_size_continuous(name = "Count", breaks = c(500, 1000, 1500, 2000)) +  # Adjust the size scale for circles
+
+  # Title and subtitles
+  labs(title = "Number of Black vs. White Individuals on Medicaid in NYC Boroughs",
+       subtitle = "Size of point indicates count, color indicates race") +
+
+  # Minimal theme and customization
+  theme_minimal() +
+  theme(legend.position = "right",
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank())
+
+```
+
+```{=tex}
+\begin{center}
+Figure 3: Map of NYC displaying number of individuals on Medicaid by Borough
+\end{center}
+```
+## Plot Description and Inferencing
+Figure 3 presents a geographical representation of New York City, displaying the number of individuals on Medicaid across different boroughs, differentiated by race.
+The boroughs—Bronx, Kings (Brooklyn), Queens, and Richmond (Staten Island)—are distinctly colored, while two circular points within each borough indicate the counts of Black/African American and White individuals.
+The size of the circles signifies the number of individuals, with larger circles representing higher counts, and the color of each circle (blue for Black/African American and red for White) represents the respective racial groups.
+A notable observation is the prominent size of the blue circle in Kings (Brooklyn), highlighting a substantial count of Black/African American individuals on Medicaid.
+Conversely, Manhattan has a noticeably larger red circle, indicating a higher number of White individuals.
+This map offers a visual synthesis of Medicaid beneficiaries' racial distribution across NYC boroughs, informing stakeholders about the demographic landscape of healthcare recipients in the metropolis.
 
@@
 
-## **Plot 2: Boxplot of Total Costs**
+## Plot 4: Waffle Plot of Severity of Illness by Race
 
@@
 #| echo: false
-cleaned_data$"Total Costs"[cleaned_data$"Total Costs" > quantile(cleaned_data$"Total Costs", 0.95)] <- quantile(cleaned_data$"Total Costs", 0.95)
+library(waffle)
+library(patchwork)
+query <- dbSendQuery(dcon,"
+SELECT Race, APRSeverityofIllnessCode FROM '2013data';
+                     ")
+IllnessCodes <- dbFetch(query, -1)
+dbClearResult(query)
 
-#summary(cleaned_data$"Total Costs")
+waffle_data1 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "Black/African American"])/2500)
+waffle_data2 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "White"])/5000)
 
-boxplot(cleaned_data$"Total Costs",
-    main = "Box Plot of Total Costs (Outliers Truncated)",
-    ylab = "Total Costs",
-    horizontal = TRUE
-)
+layout(matrix(c(1, 2), nrow = 1))
+
+plot1 <- waffle(waffle_data1, rows = 8, size = .5, 
+            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "Black/African Americans")
+plot2 <- waffle(waffle_data2, rows = 8, size = .5, 
+            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "White")
+
+arrange.plots <- plot1 + theme(plot.title = element_text(size = 10)) + plot2 + theme(plot.title = element_text(size = 10))
+arrange.plots
+```
+\begin{center}
+Figure 4: Waffle Plots Depicting Frequencies of APR Codes 
+\end{center}
+
+## Plot Description and Inferencing
+In the preceding graphical representation, we have visualized the distribution of patients based on their racial backgrounds across the various APR Severity of Illness Codes.
+These codes are categorized on a scale from 1 to 4, with higher values indicating increased severity.
+The observed trend indicates that patients of White ethnicity tend to be associated with higher severity APR codes, potentially implying a discrepancy in the prioritization of patient care between White and African American populations.
+
+\newpage
+## Plot 5: Barplots of Average Total Charges by Race 2013 vs 2021
+
+```{r, fig.height = 7, fig.width = 5}
+#| echo: false
+library(ggplot2)
+library(patchwork)
+
+pregnant_womenQ <- dbSendQuery(dcon, "
+    SELECT Race, TotalCharges
+    FROM '2013data'
+    WHERE APRMDCCode == 14 OR APRMDCCode == 15;")
+pregnant_womenDF = dbFetch(pregnant_womenQ, -1)
+dbClearResult(pregnant_womenQ)
+
+plot_race2013 <- ggplot(pregnant_womenDF) + 
+  aes(x = Race, y = TotalCharges) +
+  stat_summary(fun = "mean", geom = "bar", fill = "salmon") +  # Calculate and plot averages
+  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2013") +  # Customize axis labels
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size = 7))  # Rotate x-axis labels for
+
+pregnant_womenQ2 <- dbSendQuery(dcon, "
+    SELECT Race, TotalCharges
+    FROM '2021data'
+    WHERE APRMDCCode == 14 OR APRMDCCode == 15 AND Race != 'Multi-racial';")
+pregnant_womenDF2 = dbFetch(pregnant_womenQ2, -1)
+dbClearResult(pregnant_womenQ2)
+
+plot_race2021 <- ggplot(pregnant_womenDF2) + 
+  aes(x = Race, y = TotalCharges) +
+  stat_summary(fun = "mean", geom = "bar", fill = "violet") +  # Calculate and plot averages
+  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2021") +  # Customize axis labels
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size = 7))  # Rotate x-axis labels for
+
+combined_plot_race <- plot_race2013 + plot_race2021
+combined_plot_race
 
 ```
-<p align = "center">
-[Figure 2: Total Costs Boxplot]
-</p>
-
-### Plot Desciption and Inferencing
-
-The following boxplot refers to our Total Cost data and is representative of our data that has been cleaned to remove outliers that are in our top 5% of data points, meaning that the outliers that are extremely high are cleaned out.
-In our analysis, we used a horizontal box plot to summarize the distribution of 'Total Costs' in our dataset.
-The median, located at approximately \$8,247.96, divides the data into two halves, with the interquartile range (IQR) spanning from Q1 at \$4,519.54 to Q3 at \$15,840.81.
-Outliers, representing data points significantly deviating from the central distribution, are visible.
-The total cost range extends from \$0.41 to \$45,292.36, showcasing the data's spread.
-The mean total cost is around \$12,360.20, but outliers can influence this value.
-The boxplot also is skewed to the right.
-This visualization provides key insights into our cost structure analysis.
-
-From this we can infer that there are many outliers, however, in the discussion of costs they are heavily spread and that most patients are paying within our IQR but there are many different outliers that spread when we are discussing costs for healthcare.
-
+\begin{center}
+Figure 5: Barplots of Average Total Charges by Race in 2013 and 2021
+\end{center}
+## Plot Description and Inferencing
+In figure 5, the bar plot illustrates a compelling contrast in average total charges for hospital care among different racial groups.
+Notably, it highlights that individuals who identify as Black and those categorized under the "Other Races" group experience significantly higher average total charges compared to their White counterparts.
+This disparity in healthcare costs sheds light on the existing inequalities within the healthcare system, prompting a critical examination of the factors contributing to these disparities and the potential need for targeted interventions to address them.
+We see that in 2021, post the COVID-19
+pandemic, this disparity between average total charges of Black patients and White patients
+increased, showing that the pandemic potentially increased already present disparity in healthcare.
 \newpage
 
-## **Plot 3: Payment Type Distribution**
-
-```{r, warning = FALSE, fig.height = 15, fig.width = 50}
-#| echo: false
-# Set the margin to include more space at the bottom
-par(mar = c(5, 4, 4, 2) + 0.1)
-par(cex.main = 5 )
-cleaned_data$`Payment Typology 1` <- as.factor(cleaned_data$`Payment Typology 1`)
-payments <- table(cleaned_data$`Payment Typology 1`)
-pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA)
-legend(1, 1, 1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 3)
-
-```
-<p align = "center">
-[Figure 3: Pie Chart of Payment Distribution]
-</p>
-
-### Plot Desciption and Inferencing
-
-The following piechart identifies the payment type distribution for patients in New York in 2013.
-As we can observe, most patients are using Medicare as a way to pay for their medical bills followed by Medicaid.
-These government programs are important to New York residents and provides important data on the impact of healthcare programs.
-The exact percentagse are Medicaid at 59.7% and Medicaid at 26.7%, followed by Private Insurance at around 10%.
-This information reveals how important the impact of governmental insurance programs are to individuals in New York.
-
-\newpage
-
-## **Plot 4: Trends in Total Charges and Length of Stay**
+## Plot 6: Violin Plot Comparing New York City and Western New York's Severity of Illness
 
@@
 #| echo: false
-
-plot(cleaned_data$`Length of Stay`, cleaned_data$`Total Charges`, type = "l", col = "blue", xlab = "Length of Stay", ylab = "Total Charges", main = "Trends in Total Charges and Length of Stay")
-
-lines(cleaned_data$`Birth Weight`, cleaned_data$`Total Costs`, col = "red")
-
-legend("topright", legend = c("Total Charges", "Total Costs"), col = c("blue", "red"), lty = 1)
-
-grid()
-```
-<p align = "center">
-[Figure 4:Trends of Total Charges and Legnth of Stay ]
-</p>
-
-### Plot Desciption and Inferencing
-
-The plotted data and corresponding projections unmistakably depict a consistent total cost profile, irrespective of variations in both the length of stay and total charges.
-This apparent lack of correlation between the length of stay and associated costs prompts a deeper inquiry into the potential contributing factors.
-It would be important to explore the influence of fixed operational costs, negotiated contractual agreements, and astute cost management strategies that may explain this observed financial consistancy between the costs for a hospital and for its patients
-
-\newpage
-
-## **Plot 5: Barplot of Age Group Distribution of Patients**
-
-```{r}
-#| echo: false
-ageGroups <- table(cleaned_data$`Age Group`)
-barplot(ageGroups,
-    main = "Age Group Distribution", xlab = "Age Group",
-    ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)
-)
-
-```
-<p align = "center">
-[Figure 5: Age Group Distribution of Patients in New York]
-</p>
-
-### Plot Desciption and Inferencing
-
-Another important piece of data that we wanted to observe is the amount of patients from each age group of the population.
-The data is divided into five segmests 0-17, 18-29, 30-49, 50-69, and 70-older.
-As we can see there is an upward trend of the amount of patients as the ages of the patients increases, although there is a small decrease between 0-17 and 8-29.
-The largest frequency if the 70-older age bucket which has around 289248 compared to 18-29 with our smallest frequency of 53009.This increase in older patients reveals the impact of older populations in New York and the necesity to increase health care efforts towards older populations.
-
-\newpage
-
-## Plot 6: Stacked Bar Chart by Gender and Race
-
-```{r}
-#| echo: false
-# install.packages("ggplot2")
-library(ggplot2)
-
-plot7 <- ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
-  geom_bar() +
-  labs(
-    title = "Stacked Bar Chart by Gender and Race",
-    x = "Gender",
-    y = "Count"
-  ) +
-  scale_fill_manual(
-    values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
-    breaks = c("White", "Black", "Other"),
-    labels = c("White", "Black", "Other"),
-    name = "Race"
-  ) +
-  theme_minimal()
-
-plot7 + ggtitle("Stacked Bar Chart by Gender and Race")
-
-```
-<p align = "center">
-[Figure 6: Stacked Bar Chart by Race and Gender]
-</p>
-
-## Plot Description and Inferencing
-
-In the analysis of patient demographics in the year 2013, a bar chart was employed to visually represent the breakdown of patients by gender and further stratify each gender by race.
-A notable observation from the chart is the significant predominance of female patients over male patients during this period.
-
-Upon closer examination of the male demographic, the distribution reveals a relatively balanced representation between individuals identified as white and those classified as not white.
-This suggests a relatively even distribution among male patients concerning racial categories.
-In contrast, the female demographic exhibits a slightly more diverse distribution across racial categories.
-This observation implies that, among female patients, there is a greater degree of heterogeneity in terms of racial composition compared to their male counterparts.
-
-These findings contribute valuable insights into the gender and racial composition of the patient population in 2013, emphasizing the need for a nuanced understanding of demographic trends within healthcare settings.
-Further exploration and analysis of these demographic patterns may reveal additional insights relevant to healthcare planning, resource allocation, and the provision of patient-centered care.
-
-\newpage
-
-## Plot 7: Boxplots on the Total Costs on the Risk of Mortality
-
-```{r warning = FALSE}
-#| echo: false
-# install.packages("ggplot2")
-ggplot(cleaned_data) +
-  aes(x = `APR Risk of Mortality`, y = `Total Costs`) +
-  geom_boxplot(color = "salmon") +
-  labs(
-    title = "Total Costs on Risk of Mortality",
-    x = "APR Risk of Mortality",
-    y = "Total Costs"
-  ) +
-  theme_minimal()
-
-```
-<p align = "center">
-[Figure 7: Boxplots on Total Cost and Mortality]
-</p>
-
-## Plot Description and Inferencing
-
-In Figure 7, our objective was to assess the central tendency of total hospital costs across different levels of Risk of Mortality.
-Employing separate box plots for each severity category allowed us to discern distinct patterns in the distribution of total costs associated with varying degrees of mortality risk.
-Notably, when focusing on Extreme cases of mortality, a clear trend emerged---median total costs were significantly higher, accompanied by a broader spread in the distribution.
-This observation implies that patients facing an Extreme risk of mortality not only incur higher median costs but also exhibit greater variability in the financial implications of their healthcare, suggesting a more diverse range of medical interventions and resource utilization.
-
-This finding aligns with our initial expectation that higher levels of risk of mortality correspond to elevated costs of care.
-The discernible trend in the box plots reinforces the intuitive understanding that patients with a greater severity of illness often require more extensive and resource-intensive medical interventions, leading to increased healthcare costs.
-Beyond its immediate implications, this nuanced analysis offers valuable insights for healthcare planning and resource allocation, emphasizing the need for tailored strategies based on the specific needs associated with varying risk levels.
-
-\newpage
-
-## Plot 8: Violin Plot Comparing New York City and Western New York's Severity of Illness
-
-```{r warning = FALSE}
-#| echo: false
-subsetted_data <- cleaned_data[cleaned_data$`Health Service Area` %in% c("New York City", "Western NY"), ]
+subsetted_data <- cleaned_data[cleaned_data$`HealthServiceArea` %in% c("New York City", "Western NY"), ]
 
@@
 ```
-<p align = "center">
-[Figure 8: NYC and Western NY Severity of Illness Comparsion ]
-</p>
+
+\begin{center}
+Figure 6: NYC and Western NY Severity of Illness Comparsion
+\end{center}
 
@@
 
-Figure 8 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
+Figure 6 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY).
 The focus is on the distribution of patients' APR severity codes, measured on a 1 to 4 scale (1 being mild, 2 being moderate, 3 being serious, and 4 being extreme).
@@
 
-## Plot 9: Distribution of APR Severity of Illness for Medicaid Patients
+## Plot 7: Distribution of APR Severity of Illness for Medicaid Patients
 
@@
 ```
-<p align = "center">
-[Figure 9: APR Severity of Illness for Medicaid Patients]
-</p>
 
+\begin{center}
+Figure 7: APR Severity of Illness for Medicaid Patients
+\end{center}
 ## Plot Description and Inferencing
 
-In our examination of Figure 9, our focus shifted toward a specific subset of patients---those under Medicaid care.
+In our examination of Figure 7, our focus shifted toward a specific subset of patients---those under Medicaid care.
 Medicaid, a health program designed to support individuals classified as economically disadvantaged, prompted our investigation into the severity of diagnoses within this demographic.
@@
 
-## Plot 10: Facet Density Plots of Total Cost per Age-Group Buckets
+## Plot 8: Facet Density Plots of Total Cost per Age-Group Buckets
 
-```{r warning = FALSE}
+```{r warning = FALSE, fig.width = 7, fig.height = 5}
 #| echo: false
-gg <- ggplot(cleaned_data, aes(x = `Total Costs`, y = ..density.., fill = `Age Group`)) +
+data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)
+
+clean <- na.omit(data_set)
+
+gg <- ggplot(clean, aes(x = `Total Costs`, y = ..density.., fill = `Age Group`)) +
   geom_density(alpha = 0.5) +
@@
 
-gg + scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
+gg + 
+  scale_x_continuous(limits = c(0, 35000)) +
+  scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
   labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density (in thousands)") +
@@
 ```
-<p align = "center">
-[Figure 10: Density Plots of Total Costs per Age-Groups]
-</p>
 
+```{=tex}
+\begin{center}
+
+Figure 8: Density Plots of Total Costs per Age-Groups
+
+\end{center}
+```
 ## Plot Description and Inferencing
 
-In Figure 10, we analyzed age-specific total costs by charting density plots using a facet grid.
+In Figure 8, we analyzed age-specific total costs by charting density plots using a facet grid.
 For patients under 17, costs skewed right, suggesting lower expenses, potentially due to the overall good health of children.
@@
 
+## Plot 9: Pie Charts of Payment Type by Race
+```{r, message = FALSE}
+#| echo: false
+library(ggplot2)
+library(gridExtra)
+library(patchwork)
+
+paymentTypeAndRaceQ <- dbSendQuery(dcon, "
+                                    SELECT Race, PaymentTypology1 FROM '2013data' WHERE Race != 'Unknown';")
+paymentTypeAndRaceDF <- dbFetch(paymentTypeAndRaceQ, -1)
+dbClearResult(paymentTypeAndRaceQ)
+#str(paymentTypeAndRaceDF)
+
+pie_charts <- lapply(unique(paymentTypeAndRaceDF$Race), function(race) {
+  data_subset <- subset(paymentTypeAndRaceDF, Race == race)
+  pie_chart <- ggplot(data_subset, aes(x = "", fill = PaymentTypology1)) +
+    geom_bar(width = 1) +
+    coord_polar("y") +
+    ggtitle(race) +
+    theme_void() +
+    scale_fill_brewer(palette = "Set3") +
+    theme(legend.position = ifelse(race == 'Black/African American', "left", "none"),
+          axis.text = element_text(size = 0),
+          legend.text = element_text(size = 5)) +
+    labs(fill = "Payment Type")
+
+  return(pie_chart)
+})
+pie_charts[[1]] + pie_charts[[2]] + pie_charts[[3]] + plot_layout(ncol = 1)
+```
+
+```{=tex}
+\begin{center}
+Figure 9: Pie Charts on Payment Type by Race
+\end{center}
+```
+
+## Plot Description and Inferencing
+
+Pie Chart 1: White Patients
+
+In the first pie chart, which represents White patients, a significant portion of them use Medicare as their Payment Type.
+This choice reflects a trend among White patients, which is that a high percentage of them will rely on Medicare to cover their Hospital Charges.
+Between those that are not using Medicare, they would instead use Medicaid, Blue Cross/Blue Shield, or Private Health Insurances.
+
+Pie Chart 2: Black/African American Patients
+
+The second pie chart, which illustrates Black/African American patients, reveals a inclination towards Medicare, but a significant amount of patients would also use Medicaid.
+This suggests that there is a dual reliance on both the Medicare and Medicaid insurances, and this pie chart shows this trend.
+
+Pie Chart 3: Other Race Patients
+
+In the third pie chart, which represents the patients that are of other races, we see a greater split.
+Approximately half of the patients choose Medicare, and a majority of the patients that have not chosen Medicare rely on Medicaid instead.
+This shows a somewhat balanced usage between Medicare and Medicaid by this demographic of patients.
+
+These insights, derived from the pie charts, shed light on the nuanced payment patterns across different racial groups in the context of the year 2013, emphasizing the significance of understanding and catering to diverse payment preferences within the healthcare system of that time.
+
+\newpage
+```{r, fig.height= 6, fig.width=5}
+
+chest <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
+  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))
+
+head <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
+  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))
+
+stomach <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
+  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))
+
+chest_count <- length(chest$APRMDCCode)
+head_count <- length(head$APRMDCCode)
+stomach_count <- length(stomach$APRMDCCode)
+total_count <- length(cleaned_data$APRDRGCode)
+
+library(grid)
+library(gridExtra)
+library(scales)
+grid.newpage()
+color_palette <- colorRampPalette(c("blue", "red"))
+#Drawing the gradient key
+grad_key_height <- 0.5
+grad_key_width <- 0.05
+grad_key_x <- 0.7
+grad_key_y <- 0.5
+
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+
+grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))
+
+colors <- color_palette(100)
+for (i in 1:100) {
+  rect_gp <- gpar(fill = colors[i], col = NA)
+  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
+  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
+}
+# Add labels to the gradient key
+grid.text("Less", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
+grid.text("More", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
+
+get_color <- function(count, max_count) {
+  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
+  normalized_count <- count/max_count
+  colors <- colorRampPalette(c("blue", "purple","purple","red"))(100)  # Blue for less, Red for more
+  color_index <- max(1, ceiling(normalized_count*100))
+  return(colors[color_index])
+}
+draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+
+head_gp <- gpar(fill = "white")
+body_gp <- gpar(fill = "black")
+arms_gp <- gpar(fill = "white")
+legs_gp <- gpar(fill = "khaki")
+
+max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
+head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
+chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
+stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))
+
+draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
+draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
+draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm
+
+draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
+draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
+draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)
+
+draw_circle(0.85,.17, .02)
+draw_circle(0.85,.22, .025)
+draw_circle(0.85,.28, .03)
+draw_circle(0.85,.36, .04)
+draw_circle(0.85,.45, .05)
+draw_circle(0.85,.55, .0575)
+draw_circle(0.85,.67, .077)
+
+draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
+draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
+draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
+grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
+```
+
 ### References

### 2023-11-28 REVEALJS.qmd
Add files via upload

---
title: "Healthcare Analysis in the State of New York - Hospital Data"
subtitle: "STAT 405 Final Project" 
author: Krish Kumar, Abel Limachi, Austin Tran, Joel Villarino
format:
  revealjs: 
    transition: slide
    background-transition: fade
    transition-speed: slow
    center: true
    # embed-resources: true
    slide-number: true
    chalkboard: 
      buttons: false
    preview-links: auto
    css: styles.css
    footer: STAT 405 - Fall 2023 (Kumar, Limachi, Tran, Villarino)
server: shiny
---

# Primary Dataset Overview {background="#D2B48C"}

```{r setup, echo=FALSE,  warning = FALSE, message = FALSE, fig.align='center'}
#| context: setup

library(stringr)
library(dplyr)
library(grid)
library(gridExtra)
library(RSQLite)
library(shiny)
library(DBI)
library(RSQLite)
library(ggplot2)
library(sf)
library(waffle)
library(patchwork)
library(tigris)
library(ggplot2)
library(viridis)
library(lubridate)
library(sf)
library(DBI)
library(flextable)

dcon <- dbConnect(SQLite(), dbname = "finalDB.db", timeout = 10)
```

## 2013 New York Hospital Data

**Number of Patients**: $2,428,905$

Patient Data from the State of New York were aggregated in this dataset from the NY Health Dept. with features including:

**Variables:**

::: columns
::: {.column width="50%"}
::: {style="font-size:16px"}
-   APR Severity of Illness Code
-   APR Severity of Illness
-   Health Service Area
-   Hospital County
-   Age Group
-   Gender
-   Race
-   Ethnicity
-   Length of Stay
-   Total Charges
-   Total Costs
-   Payment Typology
-   CCS Procedure Description
-   CCS Diagnosis Code
:::
:::

::: {.column width="50%"}
::: {style="font-size:16px"}
-   CCS Diagnosis Description
-   CCS Procedure Code
-   APR MDC Code
-   APR MDC Description
-   APR DRG Code
-   APR DRG Description
-   APR Risk of Mortality
-   Abortion Edit Indicator
-   Emergency Department
-   Type of Admission
-   Facility ID
-   Facility Name
-   Zip Code (3 digit)
:::
:::
:::

## Variable Analysis

::: {style="font-size:25px"}
-   APR Severity of Illness
-   Hospital County
-   Age Group
-   Gender
-   Race
-   Ethnicity
-   Length of Stay
-   Total Charges
-   Total Costs
-   Payment Typology
:::

# Severity of Illness Visualizations {background="#D2B48C"}

## Waffle Plot of Severity of Illness by Race

```{r Waffle Plot}
query <- dbSendQuery(dcon,"
SELECT Race, APRSeverityofIllnessCode FROM '2013data';
                     ")
IllnessCodes <- dbFetch(query, -1)
dbClearResult(query)

waffle_data1 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "Black/African American"])/2480)
waffle_data2 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "White"])/5050)

layout(matrix(c(1, 2), nrow = 1))

plot1 <- waffle(waffle_data1, rows = 8, size = .5, 
            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "Black/African Americans")
plot2 <- waffle(waffle_data2, rows = 8, size = .5, 
            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "White")

arrange.plots <- plot1 + theme(plot.title = element_text(size = 10)) + plot2 + theme(plot.title = element_text(size = 10))
arrange.plots
```

## Violin Plot of Severity of Illness by Geographical Location

```{r Violin Plot}
query <- dbSendQuery(dcon,"
SELECT *
FROM '2013data';
")
cleaned_data <- dbFetch(query, -1)
dbClearResult(query)
subsetted_data <- cleaned_data[cleaned_data$`HealthServiceArea` %in% c("New York City", "Western NY"), ]

library(ggplot2)

# Create a violin plot for APR Severity of Illness Code
ggplot(subsetted_data, aes(x = factor(`HealthServiceArea`), y = `APRSeverityofIllnessCode`, fill = `HealthServiceArea`)) +
  geom_violin() +
  labs(
    x = "Health Service Area",
    y = "APR Severity of Illness Code",
    fill = "Health Service Area",
    title = "Distribution of APR Severity of Illness Code in New York City and Western NY"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Disparities in Charges and Costs {background="#D2B48C"}

## Bar Charts of Total Charge for Pregnant Women for Race (2013 & 2021)

```{r, .center}
library(ggplot2)
library(patchwork)

pregnant_womenQ <- dbSendQuery(dcon, "
    SELECT Race, TotalCharges
    FROM '2013data'
    WHERE APRMDCCode == 14 OR APRMDCCode == 15;")
pregnant_womenDF = dbFetch(pregnant_womenQ, -1)
dbClearResult(pregnant_womenQ)

plot_race2013 <- ggplot(pregnant_womenDF) + 
  aes(x = Race, y = TotalCharges) +
  stat_summary(fun = "mean", geom = "bar", fill = "salmon") +  # Calculate and plot averages
  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2013") +  # Customize axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 7))  # Rotate x-axis labels for

pregnant_womenQ2 <- dbSendQuery(dcon, "
    SELECT Race, TotalCharges
    FROM '2021data'
    WHERE APRMDCCode == 14 OR APRMDCCode == 15 AND Race != 'Multi-racial';")
pregnant_womenDF2 = dbFetch(pregnant_womenQ2, -1)
dbClearResult(pregnant_womenQ2)

plot_race2021 <- ggplot(pregnant_womenDF2) + 
  aes(x = Race, y = TotalCharges) +
  stat_summary(fun = "mean", geom = "bar", fill = "violet") +  # Calculate and plot averages
  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2021") +  # Customize axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 7))  # Rotate x-axis labels for

combined_plot_race <- plot_race2013 + plot_race2021
combined_plot_race
```

## Facet Density Plots of Total Cost per Age-Group Buckets

```{r}

gg <- ggplot(cleaned_data, aes(x = `TotalCosts`, y = ..density.., fill = `AgeGroup`)) +
  geom_density(alpha = 0.5) +
  facet_grid(`AgeGroup` ~ .) +
  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density")

gg + 
  scale_x_continuous(limits = c(0, 35000)) +
  scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density (in thousands)") +
  theme_minimal()
```

# Payment Visualizations {background="#D2B48C"}

## Pie Chart of Payment Type

```{r}
par(mar = c(5, 4, 4, 2) + 0.1)

cleaned_data$`PaymentTypology1` <- as.factor(cleaned_data$`PaymentTypology1`)
payments <- table(cleaned_data$`PaymentTypology1`)
pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA)
legend(1, 1, 1, legend = levels(cleaned_data$`PaymentTypology1`), fill = rainbow(length(payments)), cex = 0.7)

```

## Pie Charts of Payment Type by Race

```{r Pie Charts of Payment Type by Race}
library(ggplot2)
library(gridExtra)
library(patchwork)

paymentTypeAndRaceQ <- dbSendQuery(dcon, "
                                    SELECT Race, PaymentTypology1 FROM '2013data' WHERE Race != 'Unknown';")
paymentTypeAndRaceDF <- dbFetch(paymentTypeAndRaceQ, -1)
dbClearResult(paymentTypeAndRaceQ)
#str(paymentTypeAndRaceDF)

pie_charts <- lapply(unique(paymentTypeAndRaceDF$Race), function(race) {
  data_subset <- subset(paymentTypeAndRaceDF, Race == race)
  pie_chart <- ggplot(data_subset, aes(x = "", fill = PaymentTypology1)) +
    geom_bar(width = 1) +
    coord_polar("y") +
    ggtitle(race) +
    theme_void() +
    scale_fill_brewer(palette = "Set3") +
    theme(legend.position = ifelse(race == 'Black/African American', "left", "none"),
          axis.text = element_text(size = 0),
          legend.text = element_text(size = 5)) +
    labs(fill = "Payment Type")

  return(pie_chart)
})
pie_charts[[1]] + pie_charts[[2]] + pie_charts[[3]] + plot_layout(ncol = 1)
```

# Analyzing Medicaid Recipients by Race and Borough in New York City {background="#D2B48C"}

## Heat Map

```{r Heat Map}
query <- dbSendQuery(dcon, "
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
")

data <- dbFetch(query, -1)

ggplot(data, aes(x = `HospitalCounty`, y = `Race`, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(title = "Number of White and Black/Other Race Individuals on Medicaid in NYC", 
       x = "Borough", 
       y = "Race", 
       fill = "Count") +
  theme_minimal()

dbClearResult(query)

# Query the data
query <- dbSendQuery(dcon,"
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
"
)
data <- dbFetch(query, -1)

dbClearResult(query)

# Retrieve the NYC boroughs shape data and filter for the five boroughs
options(tigris_use_cache = TRUE)

# Retrieve the NYC boroughs shape data and filter for the five boroughs
ny_counties <- tigris::counties(state = "NY", cb = TRUE) %>%
  st_as_sf()

nyc_boroughs <- ny_counties[ny_counties$NAME %in% c("Bronx", "Kings", "Manhattan", "Queens", "Richmond"), ]

# Merge the data with the nyc_boroughs spatial object
data_sf <- merge(nyc_boroughs, data, by.x = "NAME", by.y = "HospitalCounty")

# Compute the centroids for the NYC boroughs for plotting circles
data_sf$centroid <- st_centroid(st_geometry(data_sf))

```

## New York City Map

```{r New York City Map}

db_path <- "finalDB.db"
con <- dbConnect(SQLite(), db_path)

# Query the data
query <- "
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
"
data <- dbGetQuery(con, query)

# Replace 'Manhattan' with 'New York' to match the shapefile data
data$HospitalCounty <- gsub("Manhattan", "New York", data$HospitalCounty)

# Disconnect from the database
dbDisconnect(con)

# Retrieve the NYC boroughs shape data and filter for the five boroughs
options(tigris_use_cache = TRUE)
ny_counties <- tigris::counties(state = "NY", cb = TRUE) %>% st_as_sf()

nyc_boroughs <- ny_counties[ny_counties$NAME %in% c("Bronx", "Kings", "New York", "Queens", "Richmond"), ]

# Merge the data with the nyc_boroughs spatial object
data_sf <- merge(nyc_boroughs, data, by.x = "NAME", by.y = "HospitalCounty", all.x = TRUE)

# Compute the centroids for the NYC boroughs for plotting circles
data_sf$centroid <- st_centroid(data_sf)

# Create a column with the X and Y coordinates of the centroids
data_sf$X <- st_coordinates(data_sf$centroid)[, "X"]
data_sf$Y <- st_coordinates(data_sf$centroid)[, "Y"]

# Plotting side-by-side maps
ggplot(data_sf) +
  geom_sf(aes(fill = NAME), color = "grey60", size = 0.5) +
  geom_point(aes(x = X, y = Y, size = count, color = Race), alpha = 0.7, show.legend = "point") +
  facet_wrap(~Race, ncol = 2) +
  scale_color_manual(values = c("Black/African American" = "blue", "White" = "red"), name = "Race") +
  scale_size_continuous(name = "Count", breaks = c(10000,20000,30000,40000,50000)) +
  labs(title = "Number of Black vs. White Individuals on Medicaid in NYC Boroughs",
       subtitle = "Size of point indicates count, color indicates county") +
  theme_minimal() +
  theme(legend.position = "right", axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())

```

# Human Body Economic Diagram (HBED) {style="font-size: 30px"}

```{r}
#| context: setup
library(shiny)
library(grid)
library(gridExtra)
library(scales)
library(dplyr)
library(RSQLite)
library(readr)
library(dplyr)
library(xtable)

dcon <- dbConnect(SQLite(), dbname = "finalDB.db", timeout = 10)
query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2021data';
  ")
  data_2021 <- dbFetch(query, -1)
  dbClearResult(query)
  query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2013data';
  ")
  data_2013 <- dbFetch(query, -1)
  dbClearResult(query)
  cleaned_data <- data_2013
```

```{r}
#| panel: sidebar
titlePanel("Interactive Human Figure")

  # Dropdown menu for selecting body parts
selectInput("bodyPart", "Select a Body Part:", choices = c("Total","Head", "Torso", "Chest"))
selectInput("time", "Select a Time Period:",
              choices = c("2013", "2021"))
```

```{r}
#| panel: fill
# Output plot created with grid graphics
plotOutput("gridPlot", width = "600px", height = "500px")
```

```{r}
#| context: server

  # This function wraps existing grid drawing code
  drawYourPlot <- function() {
    if(input$time == "2013"){
   cleaned_data <- data_2013
}
  if(input$time == "2021"){
  cleaned_data <- data_2021
}

  chest <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))

head <- select(cleaned_data, APRMDCCode, TotalCharges,LengthofStay) %>% 
  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))

stomach <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))

chest_count <- length(chest$APRMDCCode)
head_count <- length(head$APRMDCCode)
stomach_count <- length(stomach$APRMDCCode)
total_count <- length(cleaned_data$APRDRGCode)

draw_bed<- function(){
#bed frame
grid.draw(rectGrob(x = 0.80, y = 0.3, width = 0.375, height = 0.1, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
#mattress
grid.draw(rectGrob(x = 0.80, y = 0.325, width = 0.35, height = 0.12, gp = gpar(fill = "gray", col = "gray", lwd = 2)))
#pillow
grid.draw(rectGrob(x = 0.675, y = 0.38, width = 0.125, height = 0.05, gp = gpar(fill = "gray", col = "darkgray", lwd = 2)))

#left headboard
grid.draw(rectGrob(x = 0.6, y = 0.35, width = 0.05, height = 0.35, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#right headboard
grid.draw(rectGrob(x = 0.975, y = 0.325, width = 0.05, height = 0.3, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#blanket
grid.draw(rectGrob(x = 0.85, y = 0.325, width = 0.2, height = 0.135, gp = gpar(fill = "lightpink", col = "pink2", lwd = 2)))

grid.text("Average Charge per Day", x = 0.8, y = 0.5, just = "centre")

if(input$bodyPart == "Head"){
  grid.text(round(sum(head$TotalCharges)/sum(head$LengthofStay),2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Chest"){
   grid.text(round(sum(chest$TotalCharges)/sum(chest$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Torso"){
   grid.text(round(sum(stomach$TotalCharges)/sum(stomach$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
  }
}
head_gp <- gpar(fill = "white")
body_gp <- gpar(fill = "black")
arms_gp <- gpar(fill = "white")
legs_gp <- gpar(fill = "khaki")
if(input$bodyPart == "Total"){
grid.newpage()
color_palette <- colorRampPalette(c("green", "red"))
#Drawing the gradient key
grad_key_height <- 0.5
grad_key_width <- 0.05
grad_key_x <- 0.7
grad_key_y <- 0.5

draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}

grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))

colors <- color_palette(100)
for (i in 1:100) {
  rect_gp <- gpar(fill = colors[i], col = NA)
  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
}
# Add labels to the gradient key
grid.text("$", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("$$$", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("Costs", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

get_color <- function(count, max_count) {
  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
  normalized_count <- count/max_count
  colors <- colorRampPalette(c("green", "yellow", "orange", "orange", "red"))(100)  # Blue for less, Red for more
  color_index <- max(1, ceiling(normalized_count*100))
  return(colors[color_index])
}
draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}

max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))

draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm

draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)

draw_circle(0.85,.28, .03)
draw_circle(0.85,.36, .04)
draw_circle(0.85,.45, .045)
draw_circle(0.85,.555, .0525)
draw_circle(0.85,.68, .06)
grid.text("Less", x = 0.85, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("More", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("# of Cases", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
}
    # Here you would add additional grid drawing code based on the input$bodyPart
    # For example:
    else if (input$bodyPart == "Head") {
      # Add grid elements specific to the head
      grid.newpage()

grid.circle(x = 0.25, y = 0.5, r = 0.35, gp = gpar(fill = "white", col = "black", lwd = 3))
grid.lines(x = c(0.15), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))
grid.lines(x = c(0.35), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIETN FOR CIRCLE
for (i in 1:num_circles) {
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  #grid.draw()
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
  #grid.rect(x = 0.3, y = 0.25, gp = gpar(fill = "white", col = "white"), width = 2, height = .5)
  # LINES for APR Code
# 4
grid.lines(x = c(0.8, 0.9), y = c(0.7, .8), 
           gp = gpar(col = "black", lwd = 3))

grid.ellipse(x = 0.25, y = 0.85, size = 30, ar = 4, angle = 0, gp = gpar(fill = "black"))
grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()

    } else if (input$bodyPart == "Torso") {
      # Add grid elements specific to the torso
      grid.newpage()

      draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}
draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}
draw_vertical_rectangle(x = 0.45, y = 0.95, width = 0.4, height = 0.6, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is left leg
draw_vertical_rectangle(x = 0.45, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is right leg

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIETN FOR CIRCLE
for (i in 1:num_circles) { 
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
# LINES for APR Code

  # 2
grid.lines(x = c(0.75, .8), y = c(0.85, .7), 
           gp = gpar(col = "black", lwd = 3))

grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()
    } else if (input$bodyPart == "Chest") {
      # Add grid elements specific to the chest
      grid.newpage()
draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}
draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}
draw_vertical_rectangle(x = 0.4, y = 0.55, width = 0.3, height = 0.6, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.10, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is left arm
draw_vertical_rectangle(x = 0.45, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is right arm

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIENT FOR CIRCLE
for (i in 1:num_circles) { 
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
# LINES for APR Code
#3
grid.lines(x = c(0.8, .85), y = c(0.7, .85), 
           gp = gpar(col = "black", lwd = 3))

grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()
    }
  }

  # Reactive plot output based on the selection
  output$gridPlot <- renderPlot({
    drawYourPlot()
  })

chest <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))

head <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))

stomach <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))

chest_count <- length(chest$APRMDCCode)
head_count <- length(head$APRMDCCode)
stomach_count <- length(stomach$APRMDCCode)
total_count <- length(cleaned_data$APRDRGCode)
```

# Conclusion {style="font-size: 20"}

-   Disparities are examined on the basis of Race and Age
-   Between 2013 and 2021, costs and charges have significantly risen
-   Individuals who are Black/African American or from Other Races may not be receiving equitable healthcare
-   Large Amounts of Minorities Dependent on Government Programs

# Resources

2013 Dataset:

[https://www.health.ny.gov/statistics/vital_statistics/2013/](https://www.health.ny.gov/statistics/vital_statistics/2013/https://www.health.ny.gov/statistics/vital_statistics/2021/)

2021 Dataset:

[https://www.health.ny.gov/statistics/vital_statistics/2021/](https://www.health.ny.gov/statistics/vital_statistics/2013/https://www.health.ny.gov/statistics/vital_statistics/2021/)

# Questions {background="#D2B48C"}

### 2023-12-04 STAT405_FinalReport.qmd
Add files via upload

---
title: "Healthcare Insights and Sociodemographic Analysis of New York: A Comprehensive Analysis of Patient Records in 2013 and 2021"
format: pdf
author: Abel Limachi, Austin Tran, Joel Villarino, and Krish Kumar (Group 3)
editor: visual
fontsize: 10pt
geometry: margin = 1in
---

```{r, warning = FALSE, message = FALSE}
#| echo: false
library(RSQLite)
library(readr)
library(dplyr)
library(xtable)

dcon <- dbConnect(SQLite(), dbname = "finalDB.db", timeout = 10)
```

## Introduction

In the context of the ever-evolving landscape of healthcare, our choice of dataset stems from the pressing need to delve into an in-depth analysis of the heazlthcare dynamics prevailing in today's society. New York, being one of the most densely populated states in the United States, presents a unique microcosm of diverse populations and a wide spectrum of urban and rural healthcare challenges. Motivated by the significance of understanding these complexities, our research group has selected a comprehensive dataset encompassing patient records, representing healthcare interactions across various hospitals throughout New York.

The following project contains different visualizations and statistics that will provide you more information on the different social and economic factors associated with patients in the state of New York in 2013 and through 2021.

This dataset contains a wide range of healthcare-related information. It includes basic patient demographics like age, gender, race, and ethnicity; clinical data such as diagnosis and procedure codes, severity of illness, risk of mortality, and specific healthcare indicators like abortion edit and emergency department usage. Additionally, it covers financial and administrative aspects including total charges, costs, payment typology, and facility-specific information like facility ID, name, and location (represented by a 3-digit zip code). In our analysis of the dataset, we focused on specific key elements: APR Severity of Illness, Hospital County, Age Group, Gender, Race, Ethnicity, Length of Stay, Total Charges, Total Costs, and Payment Typology. These components provide a comprehensive overview of patient demographics, clinical severity, financial aspects, and hospital location, crucial for understanding healthcare dynamics and outcomes.

## Primary Dataset: Patient Records from 2013

The primary dataset that we used was obtained from the New York State Department of Health, which is the official health representative department of the state.

The dataset consists of a collection of 2,428,905 records, providing us with an extensive amount of data to work with. This rich dataset spans a spectrum of information, including patient demographics, detailed medical records, financial aspects related to their hospitalization, and critical contextual information pertinent to each patient's stay. This extensive dataset holds immense potential for generating compelling visualizations and conducting statistical analyses, offering valuable insights into the intricacies of the New York healthcare system while also providing broader insights that can contribute to the understanding of healthcare systems more generally.

## Secondary Dataset: Patient Records from 2021

The secondary dataset, like the primary one, was also obtained from the New York State Department of Health, focusing on the year 2021. This dataset had 2,493,323 records for use. The rationale for selecting this dataset is twofold:

Firstly, it allows us to track and analyze the changes in patient records across an eight-year period, offering a longitudinal perspective on the evolving healthcare landscape in New York. This longitudinal approach is crucial for understanding trends, shifts in demographic patterns, and changes in healthcare delivery and outcomes over time.

Secondly, and perhaps more importantly, the year 2021 stands as a significant historical marker due to the global COVID-19 pandemic. The pandemic has had profound and far-reaching impacts on healthcare systems worldwide, and New York was one of the epicenters of this crisis. By comparing the 2013 and 2021 datasets, we can gain invaluable insights into how the pandemic has reshaped various aspects of healthcare, including patient demographics, disease prevalence, healthcare utilization patterns, and the overall strain on the healthcare system. This comparison offers a unique opportunity to assess the resilience and adaptability of the New York healthcare system in the face of one of the most significant public health challenges of the century.

## **Exploration:**

The purpose of our exploration was to understand and contextualize the large primary dataset and investigate any trends that might exhibit potential for inequality or disparity in the healthcare system. We did this by examining a few different factors: differences in severity of illness data, disparities in charges and costs, payment typology differences, and analyzing Medicare and Medicaid recipients.

We first wanted to ground our analysis in an important part of our dataset which was exploration on patients and their reported severity of illness.

# Severity of Illness

## Plot 1: Waffle Plot of Severity of Illness by Race

```{r, fig.width = 8, fig.height = 10, message = FALSE}
#| echo: false
library(waffle)
library(patchwork)
query <- dbSendQuery(dcon,"
SELECT Race, APRSeverityofIllnessCode FROM '2013data';
                     ")
IllnessCodes <- dbFetch(query, -1)
dbClearResult(query)

waffle_data1 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "Black/African American"])/2500)
waffle_data2 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "White"])/5000)

layout(matrix(c(1, 2), nrow = 1))

plot1 <- waffle(waffle_data1, rows = 8, size = .5, 
            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "Black/African Americans")
plot2 <- waffle(waffle_data2, rows = 8, size = .5, 
            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "White")

arrange.plots <- plot1 + theme(plot.title = element_text(size = 10)) + plot2 + theme(plot.title = element_text(size = 10))
arrange.plots
```

```{=tex}
\begin{center}
Figure 1: Waffle Plots Depicting Frequencies of APR Codes 
\end{center}
```
Our first plot that we used to dive into our dataset was the above Waffle Plot, which provides an insightful visualization of the severity of illness distributed by race, focusing on Black/African American and White populations.

As we can see, the plot is divided into two sections, each representing the two racial groups mentioned. Within each section, the squares are color-coded to represent four levels of the APR severity codes, ranging from level 1 (mild, such as a common cold), indicated by the lightest color, to level 4 (severe, such as life-threatening illness), indicated by the darkest color. Across the two demographics, we observe a relatively similar distribution of APR severity codes. However, it appears that white patients, on average, has a higher concentration in severity codes level 3 and 4, which leads to these patients receiving more dire and prioritized care. As such, this raises concerns over equality: we want to ensure that patients across the board regardless of race, socioeconomic status, and other demographics receive equal care without biases.

Overall, this waffle plot serves not only as a tool to visualize data but also as a starting point for deeper discussions about health equity and the structural challenges faced by different racial groups. It's a call to action for policymakers, healthcare providers, and communities to address these disparities head-on.

With this we wanted to analyze and see if location was a potential indicator for a difference in severity of illness, this then lead us to create our next visualization which is the violin plot that compares the urban area of New York City and the rural area of Western New York.

## Plot 2: Violin Plot Comparing New York City and Western New York's Severity of Illness

```{r, fig.width = 8, warning = FALSE}
#| echo: false
query <- dbSendQuery(dcon,"
SELECT *
FROM '2013data';
")
cleaned_data <- dbFetch(query, -1)
dbClearResult(query)
subsetted_data <- cleaned_data[cleaned_data$`HealthServiceArea` %in% c("New York City", "Western NY"), ]

# Load the ggplot2 library
library(ggplot2)

# Create a violin plot for APR Severity of Illness Code
ggplot(subsetted_data, aes(x = factor(`HealthServiceArea`), y = `APRSeverityofIllnessCode`, fill = `HealthServiceArea`)) +
  geom_violin() +
  labs(
    x = "Health Service Area",
    y = "APR Severity of Illness Code",
    fill = "Health Service Area",
    title = "Distribution of APR Severity of Illness Code in Patients in New York Health Service Areas"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{=tex}
\begin{center}
Figure 2: NYC and Western NY Severity of Illness Comparsion
\end{center}
```
Figure 2 delves into a detailed breakdown of hospital patients, distinguishing between those originating from New York City (NYC) and Western New York (WNY). The focus is on the distribution of patients' APR severity codes, measured on a 1 to 4 scale (1 being mild, 2 being moderate, 3 being serious, and 4 being extreme). Notably, the violin plot reveals a higher frequency of APR severity code 2 assignments for patients from NYC compared to WNY. In the context of healthcare equality, this is a potential area of concern. In the state of New York, the areas outside of NYC are significantly more rural/suburban than NYC. Likely, the distance to the nearest hospital for residents in WNY is higher than for those who live in NYC. There being a significantly lower proportion of patients with APR severity 2 from WNY suggests that patients might not be going to the hospital unless they have severe symptoms. This could lead to many WNY residents not catching certain illnesses early when compared to NYC residents, furthering an already existing divide when it comes to healthcare access between NYC and the rest of the state..

# Disparities in Charges and Costs

## Plot 3: Barplots of Average Total Charges by Race 2013 vs 2021

```{r, fig.height = 5, fig.width = 5}
#| echo: false
library(ggplot2)
library(patchwork)

pregnant_womenQ <- dbSendQuery(dcon, "
    SELECT Race, TotalCharges
    FROM '2013data'
    WHERE APRMDCCode == 14 OR APRMDCCode == 15;")
pregnant_womenDF = dbFetch(pregnant_womenQ, -1)
dbClearResult(pregnant_womenQ)

plot_race2013 <- ggplot(pregnant_womenDF) + 
  aes(x = Race, y = TotalCharges) +
  stat_summary(fun = "mean", geom = "bar", fill = "salmon") +  # Calculate and plot averages
  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2013") +  # Customize axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 7))  # Rotate x-axis labels for

pregnant_womenQ2 <- dbSendQuery(dcon, "
    SELECT Race, TotalCharges
    FROM '2021data'
    WHERE APRMDCCode == 14 OR APRMDCCode == 15 AND Race != 'Multi-racial';")
pregnant_womenDF2 = dbFetch(pregnant_womenQ2, -1)
dbClearResult(pregnant_womenQ2)

plot_race2021 <- ggplot(pregnant_womenDF2) + 
  aes(x = Race, y = TotalCharges) +
  stat_summary(fun = "mean", geom = "bar", fill = "violet") +  # Calculate and plot averages
  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2021") +  # Customize axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 7))  # Rotate x-axis labels for

combined_plot_race <- plot_race2013 + plot_race2021
combined_plot_race

```

```{=tex}
\begin{center}
Figure 3: Barplots of Average Total Charges by Race in 2013 and 2021
\end{center}
```
In figure 3, the bar plot illustrates a compelling contrast in average total charges for hospital care among different racial groups. Notably, it highlights that individuals who identify as Black and those categorized under the "Other Races" group experience significantly higher average total charges compared to their White counterparts for their pregnancy charges. This disparity in healthcare costs sheds light on the existing inequalities within the healthcare system, prompting a critical examination of the factors contributing to these disparities and the potential need for targeted interventions to address the especially for marginalized populations of not just women on the basis of gender but also for race as well. We see that in 2021, post the COVID-19 pandemic, this disparity between average total charges of Black patients and White patients increased, showing that the pandemic potentially increased already present disparity in healthcare.

## Plot 4: Facet Density Plots of Total Cost per Age-Group Buckets

```{r warning = FALSE, fig.width = 7, fig.height = 5}
#| echo: false
data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)

clean <- na.omit(data_set)

gg <- ggplot(clean, aes(x = `Total Costs`, y = ..density.., fill = `Age Group`)) +
  geom_density(alpha = 0.5) +
  facet_grid(`Age Group` ~ .) +
  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density")

gg + 
  scale_x_continuous(limits = c(0, 35000)) +
  scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density (in thousands)") +
  theme_minimal()
```

```{=tex}
\begin{center}

Figure 4: Density Plots of Total Costs per Age-Groups

\end{center}
```
The density plots detailing Total Costs across different age groups --- 0-17, 18-29, 30-49, 50-69, and 70 or older --- illustrate a clear correlation between age and healthcare expenditure. For the younger cohorts (0-17 and 18-29), the costs are low and show little variation, reflecting the typically good health of these age groups and infrequent need for extensive healthcare. As we move into the middle-aged groups (30-49 and 50-69), the plots expand, indicating an increase in both the range and average of healthcare costs. This likely corresponds to the gradual appearance of chronic illnesses and the need for more regular medical care.

The most dispersed plot is seen in the 70+ group, signaling the highest variability and cost of healthcare. This suggests the presence of multiple, more complex health issues that are common in older age, requiring more frequent and often more expensive medical interventions.

These plots underscore the impact of aging on healthcare costs and highlight the importance of proactive health measures. There is a clear indication that early intervention and preventive care could significantly affect the healthcare system's financial demands, suggesting that investing in health education and preventive measures early on could mitigate the higher costs associated with aging populations. The analysis advocates for policy adjustments that address the escalating costs in the healthcare system, ensuring that as individuals age, they are supported by a robust and responsive healthcare infrastructure.

# Payment Visualizations

## Plot 5: Pie Charts of Payment Type

```{r, fig.height = 4, fig.width = 9}
#| echo: false
par(mar = c(5, 4, 4, 2) + 0.1)

cleaned_data$`PaymentTypology1` <- as.factor(cleaned_data$`PaymentTypology1`)
payments <- table(cleaned_data$`PaymentTypology1`)
pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA)
legend(1, 1, 1, legend = levels(cleaned_data$`PaymentTypology1`), fill = rainbow(length(payments)), cex = 0.7)

```

```{=tex}
\begin{center}
Figure 5: Pie Chart of Payment Type 
\end{center}
```
Our analysis delves into the intricacies of healthcare payments, as depicted by a comprehensive pie chart reflecting various payment types utilized by patients for hospital services. The chart initially reveals a dominance of Medicaid, Medicare, and private health insurance, indicating these as the predominant payment methods. Notably, Medicaid occupies a significant portion, underscoring its role in providing for individuals with limited financial resources. Medicare's substantial share reflects its critical support for the elderly and certain younger individuals with disabilities. The prominence of private health insurance highlights its importance in the American healthcare landscape, often associated with employment benefits.

However, the distribution also brings attention to the smaller yet significant segments such as Managed Care and Self-Pay. The Managed Care slice suggests a structured approach to healthcare, emphasizing preventive measures and cost-efficiency, while the Self-Pay category may point to gaps in healthcare coverage or a choice for those who may not want to use insurance. The presence of diverse payment sources such as Federal/State/Local/VA and the Department of Corrections indicates the healthcare system's reach, covering veterans, federal employees, and incarcerated individuals.

Building on the diverse payment types identified, we further explore the breakdown by race to uncover potential disparities. Plot 6 segments the payment types for White, Black/African American, and Other Race groups, offering a more granular view that may reflect underlying socioeconomic patterns. This breakdown by race could provide critical insights into the equity of healthcare coverage and the financial barriers different racial groups face within the healthcare system.

## Plot 6: Pie Charts of Payment Type by Race

```{r, message = FALSE}
#| echo: false
library(ggplot2)
library(gridExtra)
library(patchwork)

paymentTypeAndRaceQ <- dbSendQuery(dcon, "
                                    SELECT Race, PaymentTypology1 FROM '2013data' WHERE Race != 'Unknown';")
paymentTypeAndRaceDF <- dbFetch(paymentTypeAndRaceQ, -1)
dbClearResult(paymentTypeAndRaceQ)
#str(paymentTypeAndRaceDF)

pie_charts <- lapply(unique(paymentTypeAndRaceDF$Race), function(race) {
  data_subset <- subset(paymentTypeAndRaceDF, Race == race)
  pie_chart <- ggplot(data_subset, aes(x = "", fill = PaymentTypology1)) +
    geom_bar(width = 1) +
    coord_polar("y") +
    ggtitle(race) +
    theme_void() +
    scale_fill_brewer(palette = "Set3") +
    theme(legend.position = ifelse(race == 'Black/African American', "left", "none"),
          axis.text = element_text(size = 0),
          legend.text = element_text(size = 5)) +
    labs(fill = "Payment Type")

  return(pie_chart)
})
pie_charts[[1]] + pie_charts[[2]] + pie_charts[[3]] + plot_layout(ncol = 1)
```

```{=tex}
\begin{center}
Figure 6: Pie Charts on Payment Type by Race
\end{center}
```
The pie charts shed light on a stark disparity in healthcare payment types across racial lines. They show that White patients are more likely to use Medicare, which often correlates with a more secure financial status and access to retirement benefits. Meanwhile, Black/African American individuals and those from other racial backgrounds have a higher reliance on Medicaid, indicating lower income levels and less access to private or employer-sponsored health insurance. These trends aren't just numbers; they reflect real-world inequalities where minority groups often have less access to quality healthcare. The data doesn't just highlight a disparity; it points to the need for policies that address the economic and racial factors that limit healthcare options for some groups. It's an important piece of the puzzle for creating a more equitable healthcare system where coverage and quality care are accessible to all, regardless of racial background or economic standing.

# Analyzing Medicaid Recipients by Race and Borough in New York City

### Plot 7: New York City Medicaid Heat Map

```{r, fig.width = 10}
#| echo: false
query <- dbSendQuery(dcon, "
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
")

data <- dbFetch(query, -1)

ggplot(data, aes(x = `HospitalCounty`, y = `Race`, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(title = "Number of White and Black/Other Race Individuals on Medicaid in NYC", 
       x = "Borough", 
       y = "Race", 
       fill = "Count") +
  theme_minimal()
```

```{=tex}
\begin{center}
Figure 7: Heat Map of Patients on Medicare in  New York City Boroughs
\end{center}
```
Building on the insights from Plot 6, which highlighted the distribution of healthcare payment types among different racial groups, our focus shifted to a more localized context---New York City. This targeted analysis allows us to dissect the intricate layers of Medicaid distribution by race within the city's diverse boroughs.

In Plot 7, the heatmap provides a stark visualization of racial disparities across NYC boroughs. The variance in color intensity, from light to dark red, denotes the number of Medicaid recipients, revealing a pronounced concentration in Kings (Brooklyn), particularly among Black/African American individuals. This contrast is particularly noteworthy when compared to Manhattan, where a higher count of White individuals is evident. The heatmap's gradient effectively communicates the disparity in Medicaid dependence between these two populous boroughs. Meanwhile, Richmond (Staten Island) shows markedly lower numbers for both racial groups, hinting at different underlying socioeconomic factors at play. The Bronx and Queens present a more balanced distribution, suggesting a more equitable spread of Medicaid coverage across racial lines in these areas.

### Plot 8: New York City Map Analysis

```{r, fig.height = 6, fig.width = 7, message = FALSE, warning = FALSE}
#| echo: false
library(stringr)
library(dplyr)
library(grid)
library(gridExtra)
library(RSQLite)
library(shiny)
library(DBI)
library(RSQLite)
library(ggplot2)
library(sf)
library(waffle)
library(patchwork)
library(tigris)
library(ggplot2)
library(viridis)
library(lubridate)
library(sf)
library(DBI)
library(flextable)

db_path <- "finalDB.db"
con <- dbConnect(SQLite(), db_path)

# Query the data
query <- "
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
"
data <- dbGetQuery(con, query)

# Replace 'Manhattan' with 'New York' to match the shapefile data
data$HospitalCounty <- gsub("Manhattan", "New York", data$HospitalCounty)

# Disconnect from the database
dbDisconnect(con)

# Retrieve the NYC boroughs shape data and filter for the five boroughs
options(tigris_use_cache = TRUE)
ny_counties <- tigris::counties(state = "NY", cb = TRUE) %>% st_as_sf()

nyc_boroughs <- ny_counties[ny_counties$NAME %in% c("Bronx", "Kings", "New York", "Queens", "Richmond"), ]

# Merge the data with the nyc_boroughs spatial object
data_sf <- merge(nyc_boroughs, data, by.x = "NAME", by.y = "HospitalCounty", all.x = TRUE)

# Compute the centroids for the NYC boroughs for plotting circles
data_sf$centroid <- st_centroid(data_sf)

# Create a column with the X and Y coordinates of the centroids
data_sf$X <- st_coordinates(data_sf$centroid)[, "X"]
data_sf$Y <- st_coordinates(data_sf$centroid)[, "Y"]

# Plotting side-by-side maps
ggplot(data_sf) +
  geom_sf(aes(fill = NAME), color = "grey60", size = 0.5) +
  geom_point(aes(x = X, y = Y, size = count, color = Race), alpha = 0.7, show.legend = "point") +
  facet_wrap(~Race, ncol = 2) +
  scale_color_manual(values = c("Black/African American" = "blue", "White" = "red"), name = "Race") +
  scale_size_continuous(name = "Count", breaks = c(10000,20000,30000,40000,50000)) +
  labs(title = "Number of Black vs. White Individuals on Medicaid in NYC Boroughs",
       subtitle = "Size of point indicates count, color indicates county",
       fill = "Borough Name") + # Change the legend title for the fill aesthetic
  theme_minimal() +
  theme(legend.position = "right", axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())

```

```{=tex}
\begin{center}
Figure 8: Map of Patients on Medicare in  New York City Boroughs
\end{center}
```
Plot 8 deepens this exploration with a cartographic depiction that merges Medicaid data with the geographic layout of NYC. The boroughs are color-coded, and within each, circles of varying sizes represent the count of Medicaid recipients, with the color designating the race---blue for Black/African American and red for White. The visualization brings to light the sheer volume of Black/African American Medicaid beneficiaries in Brooklyn, as indicated by the dominance of the blue circles. In contrast, Manhattan's red circles point to a higher number of White individuals on Medicaid. This map is more than a visual aid; it encapsulates the demographic disparities in healthcare access and provides a spatial dimension to the quantitative data. For stakeholders, this map is a call to action, highlighting areas where healthcare policies and programs may need to be adjusted to address the specific needs of each borough's population, ensuring a more equitable healthcare system in one of the world's most diverse cities.

In conclusion, the sequence of plots from our analysis paints a telling picture of the healthcare payment landscape in New York City, marked by significant racial and geographic disparities. The pie chart in Plot 6 laid the foundation, showing distinct payment behaviors among racial groups. This was the impetus for a deeper dive into the local context, where the heatmap of Plot 7 highlighted stark contrasts in Medicaid distribution across boroughs, with Kings (Brooklyn) and Manhattan exhibiting pronounced disparities between Black/African American and White populations, respectively. Plot 8's map provided a vivid geographical perspective, reinforcing these disparities with visual cues that emphasize the concentration of Medicaid recipients in specific boroughs. Together, these visualizations not only underscore the existing inequities in healthcare coverage but also underscore the need for targeted policy interventions.

# Killer Plot

## Human Body Economic Diagram

# Conclusion

### 2023-12-04 STAT405_FinalReport.qmd
Add files via upload

@@
 
-```{r, fig.width = 8, warning = FALSE}
+```{r, fig.height = 4, fig.width = 8, warning = FALSE}
 #| echo: false
@@
 
-# Disparities in Charges and Costs
+# Disparities In Charges and Costs
 
-## Plot 3: Barplots of Average Total Charges by Race 2013 vs 2021
+## Plot 3: Linear Regression Model Regarding Charges and Costs
 
-```{r, fig.height = 5, fig.width = 5}
+```{r, results='asis', warning=FALSE}
+#| echo: false
+query <- dbSendQuery(dcon,"
+SELECT *
+FROM '2013data';
+")
+cleaned_data <- dbFetch(query, -1)
+dbClearResult(query)
+
+# Renaming the column to 'Length of Stay' for the linear model
+names(cleaned_data)[names(cleaned_data) == "LengthofStay"] <- "Length of Stay"
+
+y <- cleaned_data$`TotalCharges`
+length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
+
+# Set scipen to a large value to avoid scientific notation
+options(scipen = 999999999999)
+
+# Create a scatter plot
+plot(length_of_stay, y, xlab = "Length of Stay", main = "Price vs Length of Stay", ylab = "Price")
+
+# Fit the linear model
+lmfs <- lm(y ~ length_of_stay)
+
+# Add the regression line
+abline(lmfs, col = "RED")
+
+# Reset scipen to its default value after the plot
+options(scipen = 0)
+```
+
+```{=tex}
+\begin{center}
+Figure 3: Length of Stay on Price Regression Dotplot
+\end{center}
+```
+```{r, results='asis', message = FALSE, warning = FALSE, comment = FALSE}
+#| echo: false
+summary_lmfs <- summary(lmfs)
+# Assuming lmfs is your linear model
+summary_lmfs <- summary(lmfs)
+r_squared <- summary_lmfs$r.squared
+
+# Manually set the row names for the coefficients
+rownames(summary_lmfs$coefficients) <- c("Intercept", "Length of Stay")
+
+# Generate the table with xtable
+table_lmfs <- xtable::xtable(summary_lmfs$coefficients, caption = "Linear Model Summary")
+# Set column names for the table
+colnames(table_lmfs) <- c("Estimate", "Std. Error", "t value", "Pr(>|t|)")
+# Print the table w/ row names 
+print(table_lmfs, include.rownames = TRUE, comment = FALSE)
+
+```
+
+```{r, results='asis', message = FALSE, warning = FALSE}
+#| echo: false
+
+summary_lmfs <- summary(lmfs)
+
+# Extract R-squared value
+r_squared <- summary_lmfs$r.squared
+
+r_squared_df <- data.frame(`R^2` = r_squared)
+
+table_r_squared <- xtable::xtable(r_squared_df, caption = "Model R-Squared")
+
+print(table_r_squared, include.rownames = FALSE, 
+      sanitize.text.function = function(x){x}, comment = FALSE)
+
+```
+
+In our analysis of patient data for the state of New York in 2013, we conducted a linear regression to explore the impact of the length of stay on the associated price.
+The results revealed a statistically significant and strong positive relationship between these variables.
+The coefficient for length of stay that is calcualte din days was estimated at 6140.73 with a standard error of 6.622 and a remarkably high t-value of 927.39.
+This suggests that as the length of stay increases, there is a substantial increase in the price, a relationship supported by a p-value close to zero.
+The R-squared value of 0.55 indicates that approximately 55% of the price variation can be explained by the length of stay, demonstrating its relevance as a predictor.
+These findings provide important insights for healthcare management and policymaking, highlighting the need to consider and manage the cost implications of varying lengths of stay for New York state patients in 2013.
+
+However, this plot was the impetus to start seeing the interaction between total charges and costs and its relationship with other identifying factors of patients.
+
+## Plot 4: Barplots of Average Total Charges by Race 2013 vs 2021
+
+```{r, fig.height = 6, fig.width = 5}
 #| echo: false
@@
 \begin{center}
-Figure 3: Barplots of Average Total Charges by Race in 2013 and 2021
+Figure 4: Barplots of Average Total Charges by Race in 2013 and 2021
 \end{center}
 ```
-In figure 3, the bar plot illustrates a compelling contrast in average total charges for hospital care among different racial groups.
+In figure 4, the bar plot illustrates a compelling contrast in average total charges for hospital care among different racial groups.
 Notably, it highlights that individuals who identify as Black and those categorized under the "Other Races" group experience significantly higher average total charges compared to their White counterparts for their pregnancy charges.
@@
 We see that in 2021, post the COVID-19 pandemic, this disparity between average total charges of Black patients and White patients increased, showing that the pandemic potentially increased already present disparity in healthcare.
+We then wanted to start shifting our focus on looking at the age of patients.
 
-## Plot 4: Facet Density Plots of Total Cost per Age-Group Buckets
+## Plot 5: Facet Density Plots of Total Cost per Age-Group Buckets
 
@@
 
-Figure 4: Density Plots of Total Costs per Age-Groups
+Figure 5: Density Plots of Total Costs per Age-Groups
 
@@
 
-## Plot 5: Pie Charts of Payment Type
+## Plot 6: Pie Charts of Payment Type
 
@@
 \begin{center}
-Figure 5: Pie Chart of Payment Type 
+Figure 6: Pie Chart of Payment Type 
 \end{center}
@@
 
-## Plot 6: Pie Charts of Payment Type by Race
+## Plot 7: Pie Charts of Payment Type by Race
 
@@
 \begin{center}
-Figure 6: Pie Charts on Payment Type by Race
+Figure 7: Pie Charts on Payment Type by Race
 \end{center}
@@
 
-### Plot 7: New York City Medicaid Heat Map
+## Plot 8: New York City Medicaid Heat Map
 
@@
 \begin{center}
-Figure 7: Heat Map of Patients on Medicare in  New York City Boroughs
+Figure 8: Heat Map of Patients on Medicare in  New York City Boroughs
 \end{center}
@@
 
-In Plot 7, the heatmap provides a stark visualization of racial disparities across NYC boroughs.
+In Plot 8, the heatmap provides a stark visualization of racial disparities across NYC boroughs.
 The variance in color intensity, from light to dark red, denotes the number of Medicaid recipients, revealing a pronounced concentration in Kings (Brooklyn), particularly among Black/African American individuals.
@@
 
-### Plot 8: New York City Map Analysis
+## Plot 9: New York City Map Analysis
 
@@
 \begin{center}
-Figure 8: Map of Patients on Medicare in  New York City Boroughs
+Figure 9: Map of Patients on Medicare in  New York City Boroughs
 \end{center}
 ```
-Plot 8 deepens this exploration with a cartographic depiction that merges Medicaid data with the geographic layout of NYC.
+Plot 9 deepens this exploration with a cartographic depiction that merges Medicaid data with the geographic layout of NYC.
 The boroughs are color-coded, and within each, circles of varying sizes represent the count of Medicaid recipients, with the color designating the race---blue for Black/African American and red for White.
@@
 
+In our pursuit to enhance data transparency and support decision-making in healthcare, we introduce the Human Body Economic Diagram (HBED), a proposed analytical tool outlined in this report.
+The HBED represents a leap forward in medical data visualization, combining anatomical precision with economic analysis.
+It features an interactive silhouette of the human body, onto which data is mapped in an intuitive and accessible manner.
+
+The HBED's core functionality lies in its ability to visually represent medical procedure frequencies and their associated costs on the human body.
+By using variably-sized circles, each corresponding to a different body part and scaled to indicate procedural frequency, and color-coding these circles to depict the average costs, the tool provides a quick, comparative view of the medical procedures landscape.
+
+Complementing the diagram, a sophisticated dropdown menu will allow users to delve into the specifics of each anatomical region.
+Selecting a body part will reveal a detailed dashboard, presenting comprehensive metrics such as success rates of procedures, average recovery times, patient demographics, and regional cost variations.
+This granularity empowers users with the data needed to make informed decisions.
+The dropdown also includes Shiny elements to toggle between the year 2013 and 2021, and the body parts of head, chest, and torso.
+
+Designed to be responsive and versatile, the HBED tool will be accessible across a range of devices, ensuring that whether a patient seeking cost transparency, a healthcare professional comparing procedural statistics, or a policymaker analyzing economic trends in healthcare, the information is readily available.
+
 ## Human Body Economic Diagram
 
+```{r, fig.height= 3, fig.width=5, warning = FALSE, message = FALSE}
+#| echo: false
+library(shiny)
+library(grid)
+library(gridExtra)
+library(scales)
+
+query <- dbSendQuery(dcon,"
+  SELECT *
+  FROM '2021data';
+  ")
+  data_2021 <- dbFetch(query, -1)
+  dbClearResult(query)
+  query <- dbSendQuery(dcon,"
+  SELECT *
+  FROM '2013data';
+  ")
+  data_2013 <- dbFetch(query, -1)
+  dbClearResult(query)
+
+   # This function wraps your existing grid drawing code
+  drawYourPlot <- function() {
+    if(input$time == "2013"){
+  cleaned_data <- data_2013
+}
+  if(input$time == "2021"){
+  cleaned_data <- data_2021
+
+}
+
+  chest <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
+  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))
+
+head <- select(cleaned_data, APRMDCCode, TotalCharges,LengthofStay) %>% 
+  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))
+
+stomach <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
+  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))
+
+chest_count <- length(chest$APRMDCCode)
+head_count <- length(head$APRMDCCode)
+stomach_count <- length(stomach$APRMDCCode)
+total_count <- length(cleaned_data$APRDRGCode)
+
+draw_bed<- function(){
+#bed frame
+grid.draw(rectGrob(x = 0.80, y = 0.3, width = 0.375, height = 0.1, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+#mattress
+grid.draw(rectGrob(x = 0.80, y = 0.325, width = 0.35, height = 0.12, gp = gpar(fill = "gray", col = "gray", lwd = 2)))
+#pillow
+grid.draw(rectGrob(x = 0.675, y = 0.38, width = 0.125, height = 0.05, gp = gpar(fill = "gray", col = "darkgray", lwd = 2)))
+
+#left headboard
+grid.draw(rectGrob(x = 0.6, y = 0.35, width = 0.05, height = 0.35, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+
+#right headboard
+grid.draw(rectGrob(x = 0.975, y = 0.325, width = 0.05, height = 0.3, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+
+#blanket
+grid.draw(rectGrob(x = 0.85, y = 0.325, width = 0.2, height = 0.135, gp = gpar(fill = "lightpink", col = "pink2", lwd = 2)))
+
+grid.text("Average Charge per Day", x = 0.8, y = 0.5, just = "centre", gp = gpar(fontsize = 10))
+
+if(input$bodyPart == "Head"){
+  grid.text(round(sum(head$TotalCharges)/sum(head$LengthofStay),2), x = 0.85, y = 0.33, just = "centre")
+}
+else if(input$bodyPart == "Chest"){
+   grid.text(round(sum(chest$TotalCharges)/sum(chest$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
+}
+else if(input$bodyPart == "Torso"){
+   grid.text(round(sum(stomach$TotalCharges)/sum(stomach$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
+  }
+}
+head_gp <- gpar(fill = "white")
+body_gp <- gpar(fill = "black")
+arms_gp <- gpar(fill = "white")
+legs_gp <- gpar(fill = "khaki")
+if(input$bodyPart == "Total"){
+grid.newpage()
+color_palette <- colorRampPalette(c("green", "red"))
+#Drawing the gradient key
+grad_key_height <- 0.5
+grad_key_width <- 0.05
+grad_key_x <- 0.7
+grad_key_y <- 0.5
+
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+
+grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))
+
+colors <- color_palette(100)
+for (i in 1:100) {
+  rect_gp <- gpar(fill = colors[i], col = NA)
+  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
+  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
+}
+# Add labels to the gradient key
+grid.text("$", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
+grid.text("$$$", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
+grid.text("Costs", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")
+
+get_color <- function(count, max_count) {
+  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
+  normalized_count <- count/max_count
+  colors <- colorRampPalette(c("green", "yellow", "orange", "orange", "red"))(100)  # Blue for less, Red for more
+  color_index <- max(1, ceiling(normalized_count*100))
+  return(colors[color_index])
+}
+draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+
+max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
+head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
+chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
+stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))
+
+draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
+draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
+draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm
+
+draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
+draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
+draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)
+
+draw_circle(0.85,.28, .03)
+draw_circle(0.85,.36, .04)
+draw_circle(0.85,.45, .045)
+draw_circle(0.85,.555, .0525)
+draw_circle(0.85,.68, .06)
+grid.text("Less", x = 0.85, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
+grid.text("More", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
+grid.text("# of Cases", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")
+
+draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
+draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
+draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
+grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
+}
+    # Here you would add additional grid drawing code based on the input$bodyPart
+    # For example:
+    else if (input$bodyPart == "Head") {
+      # Add grid elements specific to the head
+      grid.newpage()
+
+grid.circle(x = 0.25, y = 0.5, r = 0.35, gp = gpar(fill = "white", col = "black", lwd = 3))
+grid.lines(x = c(0.15), y = c(0,0.19), 
+           gp = gpar(col = "black", lwd = 3))
+grid.lines(x = c(0.35), y = c(0,0.19), 
+           gp = gpar(col = "black", lwd = 3))
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIETN FOR CIRCLE
+for (i in 1:num_circles) {
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  #grid.draw()
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+  #grid.rect(x = 0.3, y = 0.25, gp = gpar(fill = "white", col = "white"), width = 2, height = .5)
+  # LINES for APR Code
+# 4
+grid.lines(x = c(0.8, 0.9), y = c(0.7, .8), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.ellipse(x = 0.25, y = 0.85, size = 30, ar = 4, angle = 0, gp = gpar(fill = "black"))
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+
+    } else if (input$bodyPart == "Torso") {
+      # Add grid elements specific to the torso
+      grid.newpage()
+
+      draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+draw_vertical_rectangle(x = 0.45, y = 0.95, width = 0.4, height = 0.6, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is left leg
+draw_vertical_rectangle(x = 0.45, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is right leg
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIETN FOR CIRCLE
+for (i in 1:num_circles) { 
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+# LINES for APR Code
+
+  # 2
+grid.lines(x = c(0.75, .8), y = c(0.85, .7), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+    } else if (input$bodyPart == "Chest") {
+      # Add grid elements specific to the chest
+      grid.newpage()
+draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+draw_vertical_rectangle(x = 0.4, y = 0.55, width = 0.3, height = 0.6, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.10, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is left arm
+draw_vertical_rectangle(x = 0.45, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is right arm
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIENT FOR CIRCLE
+for (i in 1:num_circles) { 
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+# LINES for APR Code
+#3
+grid.lines(x = c(0.8, .85), y = c(0.7, .85), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+    }
+  }
+
+  input <- data.frame(time  = "2013",
+                  bodyPart= "Total"
+                  )
+
+    drawYourPlot()
+
+```
+
+```{=tex}
+\begin{center}
+Figure 10: Initial Screen of HBED with Cases and Costs
+\end{center}
+```
+In Figure 10 of the Human Body Economic Diagram (HBED) for 2013, colored circles on the human silhouette represent the average cost of medical procedures for the head, torso, and chest.
+The red circle on the head indicates high-cost procedures, likely due to specialized surgeries.
+The torso's orange circle suggests moderate costs, possibly from a range of surgical interventions, while the yellow circle on the chest points to lower costs, associated with more routine procedures.
+These color-coded circles quickly convey the cost disparities of medical procedures across different body regions, encapsulating the economic aspect of healthcare in 2013.
+
+```{r, fig.height= 3, fig.width= 5, warning = FALSE, message = FALSE}
+#| echo: false
+library(shiny)
+library(grid)
+library(gridExtra)
+library(scales)
+
+query <- dbSendQuery(dcon,"
+  SELECT *
+  FROM '2021data';
+  ")
+  data_2021 <- dbFetch(query, -1)
+  dbClearResult(query)
+  query <- dbSendQuery(dcon,"
+  SELECT *
+  FROM '2013data';
+  ")
+  data_2013 <- dbFetch(query, -1)
+  dbClearResult(query)
+
+   # This function wraps your existing grid drawing code
+  drawYourPlot <- function() {
+    if(input$time == "2013"){
+  cleaned_data <- data_2013
+}
+  if(input$time == "2021"){
+  cleaned_data <- data_2021
+
+}
+
+  chest <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
+  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))
+
+head <- select(cleaned_data, APRMDCCode, TotalCharges,LengthofStay) %>% 
+  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))
+
+stomach <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
+  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))
+
+chest_count <- length(chest$APRMDCCode)
+head_count <- length(head$APRMDCCode)
+stomach_count <- length(stomach$APRMDCCode)
+total_count <- length(cleaned_data$APRDRGCode)
+
+draw_bed<- function(){
+#bed frame
+grid.draw(rectGrob(x = 0.80, y = 0.3, width = 0.375, height = 0.1, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+#mattress
+grid.draw(rectGrob(x = 0.80, y = 0.325, width = 0.35, height = 0.12, gp = gpar(fill = "gray", col = "gray", lwd = 2)))
+#pillow
+grid.draw(rectGrob(x = 0.675, y = 0.38, width = 0.125, height = 0.05, gp = gpar(fill = "gray", col = "darkgray", lwd = 2)))
+
+#left headboard
+grid.draw(rectGrob(x = 0.6, y = 0.35, width = 0.05, height = 0.35, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+
+#right headboard
+grid.draw(rectGrob(x = 0.975, y = 0.325, width = 0.05, height = 0.3, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+
+#blanket
+grid.draw(rectGrob(x = 0.85, y = 0.325, width = 0.2, height = 0.135, gp = gpar(fill = "lightpink", col = "pink2", lwd = 2)))
+
+grid.text("Average Charge per Day", x = 0.8, y = 0.5, just = "centre", gp = gpar(fontsize = 10))
+
+if(input$bodyPart == "Head"){
+  grid.text(round(sum(head$TotalCharges)/sum(head$LengthofStay),2), x = 0.85, y = 0.33, just = "centre")
+}
+else if(input$bodyPart == "Chest"){
+   grid.text(round(sum(chest$TotalCharges)/sum(chest$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
+}
+else if(input$bodyPart == "Torso"){
+   grid.text(round(sum(stomach$TotalCharges)/sum(stomach$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
+  }
+}
+head_gp <- gpar(fill = "white")
+body_gp <- gpar(fill = "black")
+arms_gp <- gpar(fill = "white")
+legs_gp <- gpar(fill = "khaki")
+if(input$bodyPart == "Total"){
+grid.newpage()
+color_palette <- colorRampPalette(c("green", "red"))
+#Drawing the gradient key
+grad_key_height <- 0.5
+grad_key_width <- 0.05
+grad_key_x <- 0.7
+grad_key_y <- 0.5
+
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+
+grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))
+
+colors <- color_palette(100)
+for (i in 1:100) {
+  rect_gp <- gpar(fill = colors[i], col = NA)
+  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
+  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
+}
+# Add labels to the gradient key
+grid.text("$", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
+grid.text("$$$", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
+grid.text("Costs", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")
+
+get_color <- function(count, max_count) {
+  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
+  normalized_count <- count/max_count
+  colors <- colorRampPalette(c("green", "yellow", "orange", "orange", "red"))(100)  # Blue for less, Red for more
+  color_index <- max(1, ceiling(normalized_count*100))
+  return(colors[color_index])
+}
+draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+
+max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
+head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
+chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
+stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))
+
+draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
+draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
+draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm
+
+draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
+draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
+draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)
+
+draw_circle(0.85,.28, .03)
+draw_circle(0.85,.36, .04)
+draw_circle(0.85,.45, .045)
+draw_circle(0.85,.555, .0525)
+draw_circle(0.85,.68, .06)
+grid.text("Less", x = 0.85, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
+grid.text("More", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
+grid.text("# of Cases", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")
+
+draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
+draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
+draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
+grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
+}
+    # Here you would add additional grid drawing code based on the input$bodyPart
+    # For example:
+    else if (input$bodyPart == "Head") {
+      # Add grid elements specific to the head
+      grid.newpage()
+
+grid.circle(x = 0.25, y = 0.5, r = 0.35, gp = gpar(fill = "white", col = "black", lwd = 3))
+grid.lines(x = c(0.15), y = c(0,0.19), 
+           gp = gpar(col = "black", lwd = 3))
+grid.lines(x = c(0.35), y = c(0,0.19), 
+           gp = gpar(col = "black", lwd = 3))
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIETN FOR CIRCLE
+for (i in 1:num_circles) {
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  #grid.draw()
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+  #grid.rect(x = 0.3, y = 0.25, gp = gpar(fill = "white", col = "white"), width = 2, height = .5)
+  # LINES for APR Code
+# 4
+grid.lines(x = c(0.8, 0.9), y = c(0.7, .8), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.ellipse(x = 0.25, y = 0.85, size = 20, ar = 4, angle = 0, gp = gpar(fill = "black"))
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+
+    } else if (input$bodyPart == "Torso") {
+      # Add grid elements specific to the torso
+      grid.newpage()
+
+      draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+draw_vertical_rectangle(x = 0.45, y = 0.95, width = 0.4, height = 0.6, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is left leg
+draw_vertical_rectangle(x = 0.45, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is right leg
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIETN FOR CIRCLE
+for (i in 1:num_circles) { 
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+# LINES for APR Code
+
+  # 2
+grid.lines(x = c(0.75, .8), y = c(0.85, .7), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+    } else if (input$bodyPart == "Chest") {
+      # Add grid elements specific to the chest
+      grid.newpage()
+draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+draw_vertical_rectangle(x = 0.4, y = 0.55, width = 0.3, height = 0.6, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.10, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is left arm
+draw_vertical_rectangle(x = 0.45, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is right arm
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIENT FOR CIRCLE
+for (i in 1:num_circles) { 
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+# LINES for APR Code
+#3
+grid.lines(x = c(0.8, .85), y = c(0.7, .85), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+    }
+  }
+
+  input <- data.frame(time  = "2013",
+                  bodyPart= "Head"
+                  )
+
+    drawYourPlot()
+
+```
+
+```{=tex}
+\begin{center}
+Figure 11: 2013 Dropdown of Head with HBED Data
+\end{center}
+```
+Figure 11 of the HBED for 2013 shows an APR Severity Level gauge by the head silhouette, highlighting high-risk and complex medical procedures, with a needle pointing towards the more severe end.
+The bar chart indicates a daily average cost of \$5,366.11, signaling the significant expenses incurred by such specialized treatments, potentially involving cutting-edge technology and high-level expertise.
+This detailed economic and clinical information aids in strategic healthcare management and the formulation of policies to ensure the delivery of cost-effective yet high-quality head medical procedures.
+Moreover, it can serve as a benchmark for tracking cost trends and evaluating the impact of new technologies or treatment protocols over time.
+
+```{r, fig.height= 3, fig.width= 5, warning = FALSE, message = FALSE}
+#| echo: false
+library(shiny)
+library(grid)
+library(gridExtra)
+library(scales)
+
+query <- dbSendQuery(dcon,"
+  SELECT *
+  FROM '2021data';
+  ")
+  data_2021 <- dbFetch(query, -1)
+  dbClearResult(query)
+  query <- dbSendQuery(dcon,"
+  SELECT *
+  FROM '2013data';
+  ")
+  data_2013 <- dbFetch(query, -1)
+  dbClearResult(query)
+
+   # This function wraps your existing grid drawing code
+  drawYourPlot <- function() {
+    if(input$time == "2013"){
+  cleaned_data <- data_2013
+}
+  if(input$time == "2021"){
+  cleaned_data <- data_2021
+
+}
+
+  chest <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
+  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))
+
+head <- select(cleaned_data, APRMDCCode, TotalCharges,LengthofStay) %>% 
+  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))
+
+stomach <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
+  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))
+
+chest_count <- length(chest$APRMDCCode)
+head_count <- length(head$APRMDCCode)
+stomach_count <- length(stomach$APRMDCCode)
+total_count <- length(cleaned_data$APRDRGCode)
+
+draw_bed<- function(){
+#bed frame
+grid.draw(rectGrob(x = 0.80, y = 0.3, width = 0.375, height = 0.1, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+#mattress
+grid.draw(rectGrob(x = 0.80, y = 0.325, width = 0.35, height = 0.12, gp = gpar(fill = "gray", col = "gray", lwd = 2)))
+#pillow
+grid.draw(rectGrob(x = 0.675, y = 0.38, width = 0.125, height = 0.05, gp = gpar(fill = "gray", col = "darkgray", lwd = 2)))
+
+#left headboard
+grid.draw(rectGrob(x = 0.6, y = 0.35, width = 0.05, height = 0.35, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+
+#right headboard
+grid.draw(rectGrob(x = 0.975, y = 0.325, width = 0.05, height = 0.3, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
+
+#blanket
+grid.draw(rectGrob(x = 0.85, y = 0.325, width = 0.2, height = 0.135, gp = gpar(fill = "lightpink", col = "pink2", lwd = 2)))
+
+grid.text("Average Charge per Day", x = 0.8, y = 0.5, just = "centre",gp = gpar(fontsize = 10))
+
+if(input$bodyPart == "Head"){
+  grid.text(round(sum(head$TotalCharges)/sum(head$LengthofStay),2), x = 0.85, y = 0.33, just = "centre")
+}
+else if(input$bodyPart == "Chest"){
+   grid.text(round(sum(chest$TotalCharges)/sum(chest$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
+}
+else if(input$bodyPart == "Torso"){
+   grid.text(round(sum(stomach$TotalCharges)/sum(stomach$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
+  }
+}
+head_gp <- gpar(fill = "white")
+body_gp <- gpar(fill = "black")
+arms_gp <- gpar(fill = "white")
+legs_gp <- gpar(fill = "khaki")
+if(input$bodyPart == "Total"){
+grid.newpage()
+color_palette <- colorRampPalette(c("green", "red"))
+#Drawing the gradient key
+grad_key_height <- 0.5
+grad_key_width <- 0.05
+grad_key_x <- 0.7
+grad_key_y <- 0.5
+
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+
+grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))
+
+colors <- color_palette(100)
+for (i in 1:100) {
+  rect_gp <- gpar(fill = colors[i], col = NA)
+  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
+  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
+}
+# Add labels to the gradient key
+grid.text("$", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
+grid.text("$$$", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
+grid.text("Costs", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")
+
+get_color <- function(count, max_count) {
+  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
+  normalized_count <- count/max_count
+  colors <- colorRampPalette(c("green", "yellow", "orange", "orange", "red"))(100)  # Blue for less, Red for more
+  color_index <- max(1, ceiling(normalized_count*100))
+  return(colors[color_index])
+}
+draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+
+max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
+head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
+chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
+stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))
+
+draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
+draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
+draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm
+
+draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
+draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
+draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)
+
+draw_circle(0.85,.28, .03)
+draw_circle(0.85,.36, .04)
+draw_circle(0.85,.45, .045)
+draw_circle(0.85,.555, .0525)
+draw_circle(0.85,.68, .06)
+grid.text("Less", x = 0.85, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
+grid.text("More", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
+grid.text("# of Cases", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")
+
+draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
+draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
+draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
+grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
+}
+    # Here you would add additional grid drawing code based on the input$bodyPart
+    # For example:
+    else if (input$bodyPart == "Head") {
+      # Add grid elements specific to the head
+      grid.newpage()
+
+grid.circle(x = 0.25, y = 0.5, r = 0.35, gp = gpar(fill = "white", col = "black", lwd = 3))
+grid.lines(x = c(0.15), y = c(0,0.19), 
+           gp = gpar(col = "black", lwd = 3))
+grid.lines(x = c(0.35), y = c(0,0.19), 
+           gp = gpar(col = "black", lwd = 3))
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIETN FOR CIRCLE
+for (i in 1:num_circles) {
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  #grid.draw()
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+  #grid.rect(x = 0.3, y = 0.25, gp = gpar(fill = "white", col = "white"), width = 2, height = .5)
+  # LINES for APR Code
+# 4
+grid.lines(x = c(0.8, 0.9), y = c(0.7, .8), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.ellipse(x = 0.25, y = 0.85, size = 20, ar = 4, angle = 0, gp = gpar(fill = "black"))
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+
+    } else if (input$bodyPart == "Torso") {
+      # Add grid elements specific to the torso
+      grid.newpage()
+
+      draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+draw_vertical_rectangle(x = 0.45, y = 0.95, width = 0.4, height = 0.6, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is left leg
+draw_vertical_rectangle(x = 0.45, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is right leg
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIETN FOR CIRCLE
+for (i in 1:num_circles) { 
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+# LINES for APR Code
+
+  # 2
+grid.lines(x = c(0.75, .8), y = c(0.85, .7), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+    } else if (input$bodyPart == "Chest") {
+      # Add grid elements specific to the chest
+      grid.newpage()
+draw_circle <- function(x, y, radius, gp = gpar()) {
+  grid.circle(x = x, y = y, r = radius, gp = gp)
+}
+draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
+  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
+}
+draw_vertical_rectangle(x = 0.4, y = 0.55, width = 0.3, height = 0.6, gp = body_gp) # this is body
+draw_vertical_rectangle(x = 0.10, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is left arm
+draw_vertical_rectangle(x = 0.45, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is right arm
+
+colors <- colorRampPalette(c("red", "yellow","green"))(1000)
+num_circles <- length(colors)
+
+# CIRCLE
+circle_radius <- 0.1
+x_center <- 0.80
+y_center <- 0.75
+
+# GRADIENT FOR CIRCLE
+for (i in 1:num_circles) { 
+  color <- colors[i]
+  angle <- seq(0, pi, length.out = num_circles + 1)[i]
+  x <- x_center + circle_radius * cos(angle) 
+  y <- y_center + circle_radius * sin(angle) 
+  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
+}
+# LINES for APR Code
+#3
+grid.lines(x = c(0.8, .85), y = c(0.7, .85), 
+           gp = gpar(col = "black", lwd = 3))
+
+grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
+draw_bed()
+    }
+  }
+
+  input <- data.frame(time  = "2021",
+                  bodyPart= "Head"
+                  )
+
+    drawYourPlot()
+
+```
+
+```{=tex}
+\begin{center}
+Figure 12: 2021 Dropdown of Head with HBED Data
+\end{center}
+```
+Transitioning to Figure 12, now shown as the 2021 Dropdown of Head with HBED Data, we see a significant change.
+The APR Severity Level remains a key feature, but the average daily charge has substantially increased to \$19,131.32.
+This stark increase could reflect several factors, such as advancements in medical technology, inflation, or changes in the healthcare market.
+
+Both dropdowns indicate that, while severity may not have dramatically changed, the economic impact on patients undergoing head-related procedures has notably increased.
+Moreover, similar dropdowns for the torso and chest are available, allowing users to explore and compare procedural frequencies, severities, and costs across different body regions over time, providing a comprehensive economic overview of medical treatments in the HBED system.
+
 # Conclusion
+
+This project, through its meticulous analysis and innovative visualization techniques, has laid bare the complex tapestry of healthcare in New York City, revealing significant disparities along racial and geographic lines.
+The data-driven insights underscore an urgent need for targeted policy interventions to address these inequities.
+
+As seen in our first two plots regarding the visualizations on the severity of illness, it becomes heavily omnipresent that there is an association with race and location.
+This adds to the discussion that there may be certain areas in the state of New York that may gain access to healthcare for more minor cases, as well as certain races and ethnicity that may not be receiving an adequate amount of care.
+
+We then worked on uncovering the disparities that exist among in the charges and costs through our linear regression model, barplots for pregnant women and our facet density plots on age.
+This revealed that on multiple identifying factors and fronts that there is a striking disparity that is observed on the basis of costs and charges.
+This inequality reveals a deep-rooted issue of economic inequality that affects multiple individuals who are vulnerable in New York that requires a more sophisticated analysis of what can be done to ensure that costs reflect a more whole and fair system.
+
+Then we presented visualizations that revealed stark contrasts in healthcare payment behaviors among different racial groups and across New York City boroughs.
+For instance, the pie chart in Plot 7 highlighted distinct payment behaviors by race, forming the basis for a more localized investigation.
+The heatmap of Plot 8 accentuated these disparities, particularly in Medicaid distribution between boroughs like Kings (Brooklyn) and Manhattan.
+Similarly, Plot 9 offered a geographical perspective, further emphasizing the concentration of Medicaid recipients and the racial disparities in healthcare access.
+
+Our project's data-driven insights reveal that individuals from minority populations often encounter significant barriers in accessing quality healthcare.
+This is not just a matter of economics but also of systemic biases and historical neglect.
+For instance, our analysis using pie charts and heatmaps clearly demonstrates how Medicaid distribution and healthcare services are unevenly spread across New York City and Western New York, or when specifically looking into New York City there is some evidence favoring some boroughs over others.
+This uneven distribution often leaves minority populations in urban areas with limited access to essential healthcare services, compared to their white counterparts.
+
+In conclusion, our project not only provides a detailed snapshot of New York's healthcare system but also serves as a call to action.
+It highlights the need for ongoing research, innovative data visualization tools, and policy reforms aimed at addressing healthcare disparities.
+By leveraging the power of data, we can pave the way for a more equitable and efficient healthcare system that caters to the needs of all New Yorkers, setting a precedent for healthcare systems globally.
+
+\newpage
+
+# Resources
+
+2013 Dataset:
+
+[https://www.health.ny.gov/statistics/vital_statistics/2013/](https://www.health.ny.gov/statistics/vital_statistics/2013/https://www.health.ny.gov/statistics/vital_statistics/2021/)
+
+2021 Dataset:
+
+[https://www.health.ny.gov/statistics/vital_statistics/2021/](https://www.health.ny.gov/statistics/vital_statistics/2013/https://www.health.ny.gov/statistics/vital_statistics/2021/)
